### 2023-08-15 Submission 1.1.tex
Homework 1

\documentclass{article}
\usepackage{enumerate}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage[table]{xcolor}
\title{Submission 1.1}
\author{Saffron Liu}
\date{}
\begin{document}
\maketitle
\begin{flushleft}
    For problems 2-12:\\
\end{flushleft}
\begin{enumerate}[(a)]
    \item \textit{Write the argument in standard form.}
    \item \textit{Claim whether the argument is valid or sound. In some cases, soundness will be difficult to determine, so “soundness is difficult to determine” is an appropriate answer.}
    \item \textit{Translate the argument into truth-functional logic. Be sure to delineate the extensions you give to atomic sentences and write the argument (now in truth-functional logic) in standard form.}
          % \item \textit{Prove that the argument is valid or invalid (as appropriate) with a proof by truth table.}
          % \item \textit{Prove that the argument is valid or invalid (as appropriate) with an informal proof.}
\end{enumerate}
\begin{flushleft}
    Problems:
\end{flushleft}
\begin{enumerate}
    \setcounter{enumi}{1}
    \item I have already said that he must have gone to King's Pyland or to Mapleton. He is not at King's Pyland, therefore he is at Mapleton. (Sir Arthur Conan Doyle) (B 20)
          \begin{enumerate}[(a)]
              \item He went to King's Pyland or to Mapleton.\\
                    He is not at King's Pyland.\\
                    \rule{15em}{.5pt}\\
                    He is at Mapleton.
              \item The argument not valid, because he might have gone to King's Pyland or Mapleton, but no longer be in either of those places. Because it is not valid, it cannot be sound.
              \item \textit{A}: He went to King's Pyland\\
                    \textit{B}: He went to Mapleton\\
                    \textit{C}: He is at King's Pyland\\
                    \textit{D}: He is at Mapleton\\\\
                    $A \lor B$\\
                    $\neg C$\\
                    \rule{5em}{.5pt}\\
                    $D$
          \end{enumerate}
    \item The patient will die unless we operate. We will operate. Therefore the patient will not die. (B 20)
          \begin{enumerate}[(a)]
              \item The patient will die unless we operate.\\
                    We will operate.\\
                    \rule{15em}{.5pt}\\
                    The patient will not die.
              \item The argument is not valid, because while operating is necessary for the patient to survive, it may not be sufficient. Since the argument is not valid, it cannot be sound.
              \item \textit{A}: We operate\\
                    \textit{B}: The patient dies\\

                    $\neg A \to B$\\
                    $A$\\
                    \rule{5em}{.5pt}\\
                    $\neg B$
          \end{enumerate}
    \item If I'm right, then I'm a fool. But if I'm a fool, I'm not right. Therefore, I'm no fool. (B 70)
          \begin{enumerate}[(a)]
              \item If I'm right, then I'm a fool.\\
                    If I'm a fool, then I'm not right.\\
                    \rule{15em}{.5pt}\\
                    I'm not a fool.
              \item The argument is not valid, since where both premises are true, it is still possible to be a fool. Since it is not valid, it cannot be sound either.
              \item \textit{A}: I'm right\\
                    \textit{B}: I'm a fool\\

                    $A \to B$\\
                    $B \to \neg A$\\
                    \rule{5em}{.5pt}\\
                    $\neg B$
          \end{enumerate}
    \item If I'm right, then I'm a fool. But if I'm a fool, I'm not right. Therefore, I'm not right. (B 70)
          \begin{enumerate}[(a)]
              \item If I'm right, then I'm a fool.\\
                    If I'm a fool, then I'm not right.\\
                    \rule{15em}{.5pt}\\
                    I'm not right.
              \item The argument is valid, but not sound (the premises are likely untrue).\\
              \textit{Saffron's Commentary: I'm not sure whether or not the premises are contradictory? On one hand, if you're right you're a fool and if you're a fool you're not right doesn't seem to work, but it is possible to be a fool and not right without violating either premise. So it is possible for them to both be true at the same time, I think? Just very strange.}
              \item \textit{A}: I'm right\\
                    \textit{B}: I'm a fool\\

                    $A \to B$\\
                    $B \to \neg A$\\
                    \rule{5em}{.5pt}\\
                    $\neg A$
          \end{enumerate}
    \item If Einstein's theory of relativity is correct, light bends in the vicinity of the sun. Light does indeed bend at the vicinity of the sun. It follows that Einstein's theory of relativity is correct. (B 70)
          \begin{enumerate}[(a)]
              \item If Einstein's theory of relativity is correct, then light bends in the vicinity of the sun.\\
                    Light bends in the vicinity of the sun.\\
                    \rule{15em}{.5pt}\\
                    Einstein's theory of relativity is correct.
              \item The argument is not valid because light could bend near the sun for a different reason than Einstein posits, so the premises would be true but the conclusion would be false. Because it is not valid, it is also not sound.
              \item \textit{A}: Einstein's theory of relativity is correct\\
                    \textit{B}: Light bends in the vicinity of the sun

                    $A \to B$\\
                    $B$\\
                    \rule{5em}{.5pt}\\
                    $A$
          \end{enumerate}
    \item Congress will agree to the cut only if the President announces his support first. The President won't announce his support first, so Congress won't agree to the cut. (B 20)
          \begin{enumerate}[(a)]
              \item Congress will agree to the cut only if the President announces his support first (If the President doesn't announce his support, Congress will not agree to the cut).\\
                    The president won't announce his support first.\\
                    \rule{15em}{.5pt}\\
                    Congress won't agree to the cut.
              \item The argument is valid and probably sound.
              \item \textit{A}: Congress will agree to the cut.\\
                    \textit{B}: The President will announce his support first.\\

                    $\neg B \to \neg A$\\
                    $\neg B$\\
                    \rule{5em}{.5pt}\\
                    $\neg A$
          \end{enumerate}
    \item If you are ambitious, you'll never achieve all your goals. But life has meaning only if you have ambition. Thus, if you achieve all your goals, life has no meaning. (B 132)
          \begin{enumerate}[(a)]
              \item If you are ambitious, you'll never achieve all your goals.\\
                    Life has meaning only if you have ambition.\\
                    \rule{15em}{.5pt}\\
                    If you achieve all your goals, then life has no meaning.
              \item The argument is valid. Soudness is difficult to determine.
              \item \textit{A}: You are ambitious.\\
                    \textit{B}: You achieve all your goals.\\
                    \textit{C}: Life has meaning.\\

                    $A \to \neg B$\\
                    $\neg A \to \neg C$\\
                    \rule{5em}{.5pt}\\
                    $B \to \neg C$
          \end{enumerate}
    \item If Adams wins the election, Brown will retire to private life. If Brown dies before the election, Adams will win it. Therefore, if Brown dies before the election, he will retire to private life. (Is this evidence that English conditionals aren't truth-functional?) (J 31)
          \begin{enumerate}[(a)]
              \item If Adams wins the election, Brown will retire to private life.\\
                    If Brown dies before the election, Adams will win the election.\\
                    \rule{15em}{.5pt}\\
                    If Brown dies before the election, he will retire to private life.
              \item The argument is valid. It is not sound because the first premise isn't always true (If Adams wins the election and Brown dies, Brown cannot retire to private life).
              \item \textit{A}: Adams wins the election.\\
                    \textit{B}: Brown retires to private life.\\
                    \textit{C}: Brown dies before the election.\\

                    $A \to B$\\
                    $C \to A$\\
                    \rule{5em}{.5pt}\\
                    $C \to B$
          \end{enumerate}
    \item "Thin is guilty," observed Watson, “because either Holmes is right and the vile Moriarty is guilty or he is wrong and the scurrilous Thin did the job; but those scoundrels are either both guilty or both innocent; and, as usual, Holmes is right.” (J 18)
          \begin{enumerate}[(a)]
              \item Holmes is right or wrong.\\
                    If Holmes is right, Moriarty is guilty.\\
                    If Holmes is wrong, Thin is guilty.\\
                    Moriarty and Thin are either both guilty or both innocent (i.e. If Moriarty is guilty, Thin is guilty).\\
                    \rule{15em}{.5pt}\\
                    Thin is guilty.
              \item The argument is valid. Soundness is difficult to determine.
              \item \textit{A}: Holmes is right.\\
                    \textit{B}: Moriarty is guilty.\\
                    \textit{C}: Thin is guilty.\\

                    $A \lor \neg A$\\
                    $A \to B$\\
                    $\neg A \to C$\\
                    $B \to C$\\
                    \rule{5em}{.5pt}\\
                    $C$
          \end{enumerate}
    \item Mittens meows exactly when she is hungry. Mittens is meowing, but she isn't hungry. Therefore the end of the Earth is at hand. (B 70)
          \begin{enumerate}[(a)]
              \item Mittens meows if and only if she is hungry.\\
                    Mittens is meowing, but she is not hungry.\\
                    \rule{15em}{.5pt}\\
                    The end of the Earth is at hand.
              \item The argument is valid but not sound because there is a contradiction in the premises.
              \item \textit{A}: Mittens meows.\\
                    \textit{B}: Mittens is hungry.\\
                    \textit{C}: The world is ending.\\

                    $A \iff B$\\
                    $A \land \neg B$\\
                    \rule{5em}{.5pt}\\
                    $C$
          \end{enumerate}
    \item God is omnipotent if and only if He can do everything. If He can't make a stone so heavy that He can't lift it, then he can't do everything. But if He can make a stone so heavy that He can't lift it, He can't do everything. Therefore, either God is not omnipotent or God does not exist. (B 132)
          \begin{enumerate}[(a)]
              \item God is omnipotent if and only if he can do everything.\\
                    If he cannot make a stone so heavy that he cannot lift it, he cannot do everything.\\
                    If he can make a stone that he cannot lift, he cannot do everything.\\
                    \rule{15em}{.5pt}\\
                    Either god is not omnipotent or he does not exist.
              \item The argument is valid. Soudness is difficult to determine.\\
                    \textit{Saffron's Commentary: I have no idea where the existence of God fits into that argument. In the English, it makes sense that it suddenly appears in the conclusion; in the logical notation, less so. I don't know whether that makes it invalid or not.}
              \item \textit{A}: God exists.\\
                    \textit{B}: God is omnipotent.\\
                    \textit{D}: God can do everything.\\
                    \textit{C}: God can make a stone which he cannot lift.\\

                    $B \iff C$\\
                    $\neg D \to \neg C$\\
                    $D \to \neg C$\\
                    \rule{5em}{.5pt}\\
                    $\neg B \lor \neg A$
          \end{enumerate}
\end{enumerate}
\begin{enumerate}
    \setcounter{enumi}{12}
    \item Consider the tic-tac-toe grid with the squares labeled as this:
          \begin{center}
              \begin{tabular}{c|c|c}
                  1 & 2 & 3 \\
                  \hline
                  4 & 5 & 6 \\
                  \hline
                  7 & 8 & 9
              \end{tabular}
          \end{center}
          Suppose that X moves first to square 5, and O moves next to square 4. Prove that X can
          guarantee a win. (Cline)\\\\
          X's next move should be to square 1.
          After these three moves, the board looks like this:
          \begin{center}
            \begin{tabular}{c|c|c}
                X & 2 & 3 \\
                \hline
                O & X & 6 \\
                \hline
                7 & 8 & 9
            \end{tabular}
        \end{center}
         At this point, if O chooses to move to any square other than square 9, X can move to square 9 on the next turn and win. See below for one possible outcome:
         \begin{center}
            \begin{tabular}{c|c|c}
                X & O & 3 \\
                \hline
                O & X & 6 \\
                \hline
                7 & 8 & X
            \end{tabular}
        \end{center}
         If O \textit{does} chose to move to square 9, X can move next to square 3. The board would then look like this:
         \begin{center}
            \begin{tabular}{c|c|c}
                X & 2 & X \\
                \hline
                O & X & 6 \\
                \hline
                7 & 8 & O
            \end{tabular}
        \end{center}
        If O played in any square other than 2, X could play in square 2 on the next turn and win. However, the same is true of square 7: if O plays in any square other than 7, X can then play in square 7 and win. 
        However, O cannot play in both squares 2 and 7 on the same turn. Regardless of where O plays, X will be able to win by playing in either square 2 or square 7.

\end{enumerate}
\end{document}

### 2023-08-16 Submission 1.1.tex
Update questions 4 and 5 - Right -> Fool -> Not Right is a contradiction - Contradiction in premises means it's possible to prove A -> !A; NOT that there's no case in which they're all true

@@
                     I'm not a fool.
-              \item The argument is not valid, since where both premises are true, it is still possible to be a fool. Since it is not valid, it cannot be sound either.
+              \item The argument is neither valid nor sound because the premises form a contradiction.
+Being right implies being a fool, but being a fool implies being not right, so being right implies being not right.
               \item \textit{A}: I'm right\\
@@
                     I'm not right.
-              \item The argument is valid, but not sound (the premises are likely untrue).\\
+              \item Again, the premises form a contradiction and so the argument can't be either valid or sound.\\  
               \textit{Saffron's Commentary: I'm not sure whether or not the premises are contradictory?
@@
 So it is possible for them to both be true at the same time, I think?
-Just very strange.}
+Just very strange.\\
+              Update: Contradictory means you can prove $A \to \neg A$, regardless of whether or not there is a case where all premises are true.}
               \item \textit{A}: I'm right\\

### 2023-08-16 Submission 1.1.tex
Possible contradiction about question 12 (God) - Wanted to figure out how to incorporate the existence of God - Now unresolved - There's a contradiction, but it's the same one as is present in Descartes' cogito

@@
 In the English, it makes sense that it suddenly appears in the conclusion; in the logical notation, less so.
-I don't know whether that makes it invalid or not.}
+I don't know whether that makes it invalid or not.\\
+                    Update: It seems like there's a contradiction posed by $\neg D \to \neg C$ and $D \to \neg C$, because the second statemnent is equivalent to $C \to \neg D$, meaning $C \to \neg C$.
+However, Descartes'} cogito \textit{has the same issue.}
               \item \textit{A}: God exists.\\
                     \textit{B}: God is omnipotent.\\
-                    \textit{D}: God can do everything.\\
-                    \textit{C}: God can make a stone which he cannot lift.\\
+                    \textit{C}: God can do everything.\\
+                    \textit{D}: God can make a stone which he cannot lift.\\
 
-                    $B \iff C$\\
+                    $B \leftrightarrow C$\\
                     $\neg D \to \neg C$\\

### 2023-08-19 Submission 1.1.tex
Homework 2 - Knights and Knaves - Fixed validity on some problems (contradictory -> valid but not sound) - Truth tables (part d)

@@
 \usepackage[table]{xcolor}
+
 \title{Submission 1.1}
@@
 \maketitle
+
+\begin{enumerate}
+      \item Knaves always lie, knights always tell the truth, and in Transylvania, where everybody is one or the other (but you can’t tell which by looking), you encounter two people, one of whom says, “He’s a knight or I’m a knave.” What are they?
+(J 20)\\\\
+            A: I'm a knight.\\
+            B: He's a knight.\\
+            C: He's a knight or I'm a knave.\\
+            Notes:\\
+            If being a knight is false, being a knave is true.\\
+            If I am a knight, then (he's a knight or I'm not a knight). If I'm not a knight, then not (he's a knight or I'm not a knight).\\
+            To solve the problem, we must find a combination of knight and knave such that the truth value of "he's a knight or I'm a knave" is consistent with whether or not I'm a knight.
+Therefore, we're searching for a value where C and A are equivalent.\\
+            \begin{tabular}{c|c|c|c}
+                  A & B & $C: B \lor \neg A$ & $A \leftrightarrow C$ \\
+                  \hline
+                  T & T & T                  & T                     \\
+                  T & F & F                  & F                     \\
+                  F & T & T                  & F                     \\
+                  F & F & T                  & F                     \\
+            \end{tabular}
+            Both he and I are knights.
+\end{enumerate}
 \begin{flushleft}
@@
 Be sure to delineate the extensions you give to atomic sentences and write the argument (now in truth-functional logic) in standard form.}
-          % \item \textit{Prove that the argument is valid or invalid (as appropriate) with a proof by truth table.}
+      \item \textit{Prove that the argument is valid or invalid (as appropriate) with a proof by truth table.}
             % \item \textit{Prove that the argument is valid or invalid (as appropriate) with an informal proof.}
 \end{enumerate}
-\begin{flushleft}
-    Problems:
-\end{flushleft}
 \begin{enumerate}
@@
                         $\neg B$
+                  \item
+                        \begin{tabular}{c|c|c}
+                              A                      & B & $\neg A \to B$ \\
+                              \hline
+                              T                      & T & T              \\
+                              T                      & F & T              \\
+                              \rowcolor{lightgray} F & T & T              \\
+                              \rowcolor{lightgray} F & F & F
+                        \end{tabular} \\
+                        Where both premises are true, B can still be either true or false. This means there is no guarantee the patient will not die, so the argument is invalid.
             \end{enumerate}
@@
                         I'm not a fool.
-              \item The argument is neither valid nor sound because the premises form a contradiction.
+                  \item The argument is valid but not sound because the premises form a contradiction.
 Being right implies being a fool, but being a fool implies being not right, so being right implies being not right.
@@
                         I'm not right.
-              \item Again, the premises form a contradiction and so the argument can't be either valid or sound.\\  
+                  \item Again, the premises form a contradiction and so the argument is valid but not sound.\\
                         \textit{Saffron's Commentary: I'm not sure whether or not the premises are contradictory?
@@
                         $\neg A$
+                  \item
+                        \begin{tabular}{c|c|c|c}
+                              A                      & B & $A \to B$ & $B \to \neg A$ \\
+                              \hline
+                              \rowcolor{lightgray} T & T & T         & F              \\
+                              \rowcolor{lightgray} T & F & F         & T              \\
+                              F                      & T & T         & T              \\
+                              F                      & F & T         & T              \\
+                        \end{tabular} \\
+                        The argument is valid because in all rows where the premises are met, I am not right.\\
+                        \textit{Saffron's Commentary: I'm not actually sure how to represent a contradiction in a truth table.
+It would work if there were no case where the contradictory statements could be true at the same time, but that's not the case here.}
             \end{enumerate}
@@
                         $\neg A$
+                  \item \begin{tabular}{c|c|c}
+                              A                      & B & $(\neg B \to \neg A) = (A \to B)$ \\
+                              \hline
+                              \rowcolor{lightgray} T & T & T                                 \\
+                              \rowcolor{lightgray} T & F & F                                 \\
+                              \rowcolor{lightgray} F & T & T                                 \\
+                              F                      & F & T
+                        \end{tabular} \\ The argument is valid because the only row where both premises are true is where Congress does not agree to the cut.
             \end{enumerate}
@@
                         $C \to B$
+                  \item
+                        \begin{tabular}{c|c|c|c|c}
+                              A                      & B & C & $A \to B$ & $C \to A$ \\
+                              T                      & T & T & T         & T         \\
+                              T                      & T & F & T         & T         \\
+                              \rowcolor{lightgray} T & F & T & F         & T         \\
+                              \rowcolor{lightgray} T & F & F & F         & T         \\
+                              \rowcolor{lightgray} F & T & T & T         & F         \\
+                              F                      & T & F & T         & T         \\
+                              \rowcolor{lightgray} F & F & T & T         & F         \\
+                              F                      & F & F & T         & T         \\
+                        \end{tabular} \\ In the rows where both premises are met, there is only one case where Brown dies before the election, and in this case he does indeed retire to private life.
+This means the argument is valid.
             \end{enumerate}
@@
                         $C$
+                  \item \begin{tabular}{c|c|c|c|c}
+                              A & B & C & $A \iff B$ & $A \land \neg B$ \\
+                              \hline
+                              \rowcolor{lightgray} T & T & T & T          & F                \\
+                              \rowcolor{lightgray} T & T & F & T          & F                \\
+                              \rowcolor{lightgray} T & F & T & F          & T                \\
+                              \rowcolor{lightgray} T & F & F & F          & T                \\
+                              \rowcolor{lightgray} F & T & T & F          & F                \\
+                              \rowcolor{lightgray} F & T & F & F          & F                \\
+                              \rowcolor{lightgray} F & F & T & T          & F                \\
+                              \rowcolor{lightgray} F & F & F & T          & F                \\
+                        \end{tabular}\\
+                        There is no case where all premises are true, so the premises are contradictory and thus the argument is valid.
             \end{enumerate}

### 2023-08-21 Submission 1.1.tex
Problems 4&5 contradiction

@@
                         I'm not a fool.
-                  \item The argument is valid but not sound because the premises form a contradiction.
-Being right implies being a fool, but being a fool implies being not right, so being right implies being not right.
+                  \item The argument is invalid and unsound. If you look at the truth table in 5d, you can see that there is a counterexample where I'm not right and also a fool.
                   \item \textit{A}: I'm right\\
@@
                         I'm not right.
-                  \item Again, the premises form a contradiction and so the argument is valid but not sound.\\
+                  \item The argument is valid, because being right poses a contradiction with the premises, but being not right doesn't cause any contradiction. Therefore, I must not be right\\
                         \textit{Saffron's Commentary: I'm not sure whether or not the premises are contradictory?
@@
 Just very strange.\\
-                              Update: Contradictory means you can prove $A \to \neg A$, regardless of whether or not there is a case where all premises are true.}
+                              Update: An argument is always valid where the premises are inconsistent, because then there is never a counterexample.}
                   \item \textit{A}: I'm right\\
@@
                         The argument is valid because in all rows where the premises are met, I am not right.\\
-                        \textit{Saffron's Commentary: I'm not actually sure how to represent a contradiction in a truth table.
-It would work if there were no case where the contradictory statements could be true at the same time, but that's not the case here.}
             \end{enumerate}

### 2023-08-23 Submission 1.1.tex
HW 3 - 2-7e - 12e + update to it  (God problem) - 14-20

@@
 \begin{enumerate}
-      \item Knaves always lie, knights always tell the truth, and in Transylvania, where everybody is one or the other (but you can’t tell which by looking), you encounter two people, one of whom says, “He’s a knight or I’m a knave.” What are they?
+      \item Knaves always lie, knights always tell the truth, and in Transylvania, where everybody is one or the other (but you can't tell which by looking), you encounter two people, one of whom says, “He's a knight or I'm a knave.” What are they?
 (J 20)\\\\
@@
 \end{enumerate}
+
 \begin{flushleft}
@@
       \item \textit{Prove that the argument is valid or invalid (as appropriate) with a proof by truth table.}
-            % \item \textit{Prove that the argument is valid or invalid (as appropriate) with an informal proof.}
+      \item \textit{Prove that the argument is valid or invalid (as appropriate) with an informal proof.}
 \end{enumerate}
+
 \begin{enumerate}
@@
                         He is at Mapleton.
-                  \item The argument not valid, because he might have gone to King's Pyland or Mapleton, but no longer be in either of those places. Because it is not valid, it cannot be sound.
+                  \item The argument invalid and unsound.
                   \item \textit{A}: He went to King's Pyland\\
@@
                         $D$
+                        \setcounter{enumii}{4}
+                  \item The argument is invalid because he might have gone to King's Pyland or Mapleton, but no longer be in either of those places.
+For example, he may have gone to Mapleton, but since left and thus no longer be there.
             \end{enumerate}
@@
                         The patient will not die.
-                  \item The argument is not valid, because while operating is necessary for the patient to survive, it may not be sufficient. Since the argument is not valid, it cannot be sound.
+                  \item The argument is invalid and unsound.
                   \item \textit{A}: We operate\\
@@
                         Where both premises are true, B can still be either true or false. This means there is no guarantee the patient will not die, so the argument is invalid.
+                  \item The argument is not valid, because while operating is necessary for the patient to survive, it may not be sufficient.
             \end{enumerate}
@@
                         I'm not a fool.
-                  \item The argument is invalid and unsound. If you look at the truth table in 5d, you can see that there is a counterexample where I'm not right and also a fool.
+                  \item The argument is invalid and unsound.
                   \item \textit{A}: I'm right\\
@@
                         $\neg B$
+                        \setcounter{enumii}{4}
+                  \item  It is possible to be not right and also a fool.
             \end{enumerate}
@@
                         I'm not right.
-                  \item The argument is valid, because being right poses a contradiction with the premises, but being not right doesn't cause any contradiction. Therefore, I must not be right\\
+                  \item The argument is valid, but unsound because the premises may not be true.\\
                         \textit{Saffron's Commentary: I'm not sure whether or not the premises are contradictory?
@@
                         The argument is valid because in all rows where the premises are met, I am not right.\\
+                  \item The argument is valid.
+If I assume I am right, then I am a fool, and then I'm not right.
+My assumption creates a contradiction, therefore it must be false and I must not be right.
             \end{enumerate}
@@
                         Einstein's theory of relativity is correct.
-                  \item The argument is not valid because light could bend near the sun for a different reason than Einstein posits, so the premises would be true but the conclusion would be false.
-Because it is not valid, it is also not sound.
+                  \item The argument is invalid and unsound.
                   \item \textit{A}: Einstein's theory of relativity is correct\\
@@
                         $A$
+                        \setcounter{enumii}{4}
+                  \item The argument is not valid because light could bend near the sun for a different reason than Einstein posits, so the premises would be true but the conclusion would be false.
             \end{enumerate}
@@
                         \end{tabular} \\ The argument is valid because the only row where both premises are true is where Congress does not agree to the cut.
+                  \item The president will not announce his support.
+If the president does not announce support, Congress will not agree to the cut.
+Therefore, Congress will not agree to the cut and the argument is valid.
             \end{enumerate}
@@
 I don't know whether that makes it invalid or not.\\
-                              Update: It seems like there's a contradiction posed by $\neg D \to \neg C$ and $D \to \neg C$, because the second statemnent is equivalent to $C \to \neg D$, meaning $C \to \neg C$.
-However, Descartes'} cogito \textit{has the same issue.}
+                              Update: I'm choosing to make his existence biconditional with his omnipotence.}
                   \item \textit{A}: God exists.\\
@@
 
+                        $A \leftrightarrow B$\\
                         $B \leftrightarrow C$\\
@@
                         $\neg B \lor \neg A$
+                  \item Either God can or cannot make a stone so heavy that he cannot lift it.
+In either case, he cannot do anything and is thus not omnipotent.
+Thus, he is either not omnipotent, or does not exist, as his existence is biconditional with his omnipotence.
             \end{enumerate}
 \end{enumerate}
+
+\begin{flushleft}
+      For problems 13-20:
+\end{flushleft}
+
+\begin{itemize}
+      \item \textit{If the claim that's made is correct, prove that it's correct.}
+      \item \textit{If the claim that's made is incorrect, prove that it's incorrect.}
+      \item \textit{If the problem asks a question, answer it with a correct claim, and prove that your
+                  claim is correct.}
+\end{itemize}
+
 \begin{enumerate}
@@
             However, O cannot play in both squares 2 and 7 on the same turn. Regardless of where O plays, X will be able to win by playing in either square 2 or square 7.
+      \item You can't make a valid argument invalid by adding premises. (J 19)\\
+            This claim is correct.
+An argument is invalid if and only if there is a counterexample, i.e.
+a case where all premises are true but the conclusion is false.
+Adding a premise can never cause a case where not all premises were true to become a case with all true premises.
+Thus, adding a premise can never increase the number of counterexamples an argument has.
+Because valid arguments have no counterexamples, adding a premise to a valid argument will keep the argument's state as having no counterexamples.
+Thus, the argument will still be valid.
+      \item You can't make an invalid argument valid by removing premises. (J 19)\\
+            This claim is correct.
+An argument is invalid if and only if it has one or more counterexample.
+Removing premises from an argument cannot cause a case to go from having all true premises to having at least one untrue premise.
+Thus, removing premises can only create new counterexamples, not remove counterexamples.
+Therefore, removing premises from an argument with at least one counterexample can never cause the argument to have no counterexamples, so removing premises from an invalid argument can never create a valid argument.
+      \item Suppose ($A \land B) \to C$ is contingent. What can you say about the argument \begin{tabular}{c}A\\B\\\hline C\end{tabular}? (M 46)\\
+            \textit{Saffron's Commentary: I don't really know what contingent means!
+It's defined in the textbook, but I don't know if this question means "C is contingent upon $A \land B$" or simply "$(A \land B) \to C$" may be true or untrue depending on the case.
+I'm assuming the latter for this problem, but who knows.
+                  Update: it's fine, probably? See theorem-argument exchange (Theorem 1.1).}\\
+            The argument is invalid.
+For ($A \land B) \to C$ to be contingent (and not a tautology), there must be at least one case where the conditional is false.
+This means there is a case where $A \land B$ is true but C is false, which would be a counterexample to the argument, rendering it invalid.
+      \item Suppose the argument \begin{tabular}{c}A\\B\\\hline C\end{tabular} is valid. What can you say about ($A \land B) \to C$?\\
+            ($A \land B) \to C$ is a tautology.
+Because the argument is valid, there is no case where all premises (both A and B) are true but C is false.
+Such a case would be the only one in which the conditional was not true, but because it does not exist, ($A \land B) \to C$ must always be true.
+      \item Some arguments with contradictory premises aren't valid. (B 59)\\
+            This claim is incorrect.
+For an argument to be valid, there must be a counterexample.
+However, if the premises are contradictory, there is no case where all premises are true, and thus no counterexamples.
+      \item Suppose that \{A, B, C\} is inconsistent. What can you say about $A \land B \land C$? (M 46)\\
+            $A \land B \land C$ is always false and therefore a contradiction.
+Because \{A, B, C\} is inconsistent, there is no case in which they are all true, so there must always be at least one atomic sentence which is false.
+$A \land B \land C$ is false if at least one of it's atomic sentences is false.
+      \item Some arguments whose conclusions are contradictions are valid.\\
+            This claim is true. Any argument with inconsistent premises will never have a counterexample, because there is no case where all premises are true. Thus, the argument would be valid.
 

### 2023-08-23 Submission 1.1.tex
Fixed quote marks

@@
             If I am a knight, then (he's a knight or I'm not a knight). If I'm not a knight, then not (he's a knight or I'm not a knight).\\
-            To solve the problem, we must find a combination of knight and knave such that the truth value of "he's a knight or I'm a knave" is consistent with whether or not I'm a knight.
+            To solve the problem, we must find a combination of knight and knave such that the truth value of “he's a knight or I'm a knave” is consistent with whether or not I'm a knight.
 Therefore, we're searching for a value where C and A are equivalent.\\
@@
             \end{enumerate}
-      \item "Thin is guilty," observed Watson, “because either Holmes is right and the vile Moriarty is guilty or he is wrong and the scurrilous Thin did the job; but those scoundrels are either both guilty or both innocent; and, as usual, Holmes is right.” (J 18)
+      \item “Thin is guilty,” observed Watson, “because either Holmes is right and the vile Moriarty is guilty or he is wrong and the scurrilous Thin did the job; but those scoundrels are either both guilty or both innocent; and, as usual, Holmes is right.” (J 18)
             \begin{enumerate}[(a)]
@@
             \textit{Saffron's Commentary: I don't really know what contingent means!
-It's defined in the textbook, but I don't know if this question means "C is contingent upon $A \land B$" or simply "$(A \land B) \to C$" may be true or untrue depending on the case.
+It's defined in the textbook, but I don't know if this question means “C is contingent upon $A \land B$” or simply “$(A \land B) \to C$” may be true or untrue depending on the case.
 I'm assuming the latter for this problem, but who knows.

### 2023-08-25 Submission 1.1.tex
Polishing - Minor formatting changes - Clarifying explanations

@@
                   T & T & T                  & T                     \\
-                  T & F & F                  & F                     \\
-                  F & T & T                  & F                     \\
-                  F & F & T                  & F                     \\
+                  \rowcolor{lightgray} T & F & F                  & F                     \\
+                  \rowcolor{lightgray} F & T & T                  & F                     \\
+                  \rowcolor{lightgray} F & F & T                  & F                     \\
             \end{tabular}
@@
                         If you achieve all your goals, then life has no meaning.
-                  \item The argument is valid. Soudness is difficult to determine.
+                  \item The argument is valid. If you achieve all your goals, you are not ambitious, and if you are not ambitious, life has no meaning. Soudness is difficult to determine.
                   \item \textit{A}: You are ambitious.\\
@@
                         $\neg B \lor \neg A$
+                  \setcounter{enumii}{4}
                   \item Either God can or cannot make a stone so heavy that he cannot lift it.
@@
 However, the same is true of square 7: if O plays in any square other than 7, X can then play in square 7 and win.
-            However, O cannot play in both squares 2 and 7 on the same turn. Regardless of where O plays, X will be able to win by playing in either square 2 or square 7.
+            However, O cannot play in both squares 2 and 7 on the same turn.
+Regardless of where O plays, X will be able to win by playing in either square 2 or square 7.
+Therefore, regardless of where O plays on any turn, X is able to guarantee a win.
       \item You can't make a valid argument invalid by adding premises. (J 19)\\
@@
       \item Some arguments whose conclusions are contradictions are valid.\\
-            This claim is true. Any argument with inconsistent premises will never have a counterexample, because there is no case where all premises are true. Thus, the argument would be valid.
+            This claim is correct.
+Any argument with inconsistent premises will never have a counterexample, because there is no case where all premises are true.
+Thus an argument with inconsistent premises and a contradicion for a conclusion would be valid.
 

### 2023-08-25 Submission 1.1.tex
Remove name + commentary Fixed wording in God problem

@@
 \title{Submission 1.1}
-\author{Saffron Liu}
+% \author{Saffron Liu}
 \date{}
@@
                   \item The argument is valid, but unsound because the premises may not be true.\\
-                        \textit{Saffron's Commentary: I'm not sure whether or not the premises are contradictory?
+                        % \textit{Saffron's Commentary: I'm not sure whether or not the premises are contradictory?
 On one hand, if you're right you're a fool and if you're a fool you're not right doesn't seem to work, but it is possible to be a fool and not right without violating either premise.
@@
 Just very strange.\\
-                              Update: An argument is always valid where the premises are inconsistent, because then there is never a counterexample.}
+                        %       Update: An argument is always valid where the premises are inconsistent, because then there is never a counterexample.}
                   \item \textit{A}: I'm right\\
@@
                   \item The argument is valid. Soudness is difficult to determine.\\
-                        \textit{Saffron's Commentary: I have no idea where the existence of God fits into that argument.
+                        % \textit{Saffron's Commentary: I have no idea where the existence of God fits into that argument.
 In the English, it makes sense that it suddenly appears in the conclusion; in the logical notation, less so.
 I don't know whether that makes it invalid or not.\\
-                              Update: I'm choosing to make his existence biconditional with his omnipotence.}
+                        %       Update: I'm choosing to make his existence biconditional with his omnipotence.}
                   \item \textit{A}: God exists.\\
@@
                   \item Either God can or cannot make a stone so heavy that he cannot lift it.
-In either case, he cannot do anything and is thus not omnipotent.
+In either case, he cannot do everything and is thus not omnipotent.
 Thus, he is either not omnipotent, or does not exist, as his existence is biconditional with his omnipotence.
@@
       \item Suppose ($A \land B) \to C$ is contingent. What can you say about the argument \begin{tabular}{c}A\\B\\\hline C\end{tabular}? (M 46)\\
-            \textit{Saffron's Commentary: I don't really know what contingent means!
+            % \textit{Saffron's Commentary: I don't really know what contingent means!
 It's defined in the textbook, but I don't know if this question means “C is contingent upon $A \land B$” or simply “$(A \land B) \to C$” may be true or untrue depending on the case.
 I'm assuming the latter for this problem, but who knows.
-                  Update: it's fine, probably? See theorem-argument exchange (Theorem 1.1).}\\
+            %       Update: it's fine, probably? See theorem-argument exchange (Theorem 1.1).}\\
             The argument is invalid.

### 2023-08-30 Submission 1.2.tex
HW 4 - 10-13

\documentclass{article}
\usepackage{enumerate}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage[table]{xcolor}

\title{Submission 1.2}
\author{Saffron Liu}
\date{}
\begin{document}
\maketitle

\begin{flushleft}
    For problems 10-13:
\end{flushleft}
\begin{itemize}
    \item \textit{If the claim that’s made is correct, prove that it’s correct.}
    \item \textit{If the claim that’s made is incorrect, prove that it’s incorrect.}
    \item \textit{If the problem asks a question, answer it with a correct claim, and prove that your claim is correct.}
\end{itemize}

\begin{enumerate}
    \setcounter{enumi}{9}
    \item A two-place connective, $\circ$, is called associative if $(A \circ B) \circ C$ is logically equivalent to $A \circ (B \circ C)$. Which of $\land, \lor, \to, \leftrightarrow$ are associative? (J 20)\\
          $\land, \lor$, and $\leftrightarrow$ are associative. $\to$ is not.\\
          $(A \land B) \land C$ and $A \land (B \land C)$ are logically equivalent because each is true if and only if A, B, and C are all true; if at least one is false, the expression is as well.\\
          $(A \lor B) \lor C$ and $A \lor (B \lor C)$ are logically equivalent because each is false if and only if A, B, and C are all false; if at least one is true, the expression is true.\\
          $(A \leftrightarrow B) \leftrightarrow C$ and $A \leftrightarrow (B \leftrightarrow C)$ are logically equivalent. I can't explain it, but hey look, I made a truth table!
          \begin{tabular}{c|c|c|c|c}
              A & B & C & $(A \leftrightarrow B) \leftrightarrow C$ & $A \leftrightarrow (B \leftrightarrow C)$ \\
              \hline
              T & T & T & T                                         & T                                         \\
              T & T & F & F                                         & F                                         \\
              T & F & T & F                                         & F                                         \\
              T & F & F & T                                         & T                                         \\
              F & T & T & F                                         & F                                         \\
              F & T & F & T                                         & T                                         \\
              F & F & T & T                                         & T                                         \\
              F & F & F & F                                         & F
          \end{tabular}\\
          % $(A \leftrightarrow B) \leftrightarrow C$ and $A \leftrightarrow (B \leftrightarrow C)$ are logically equivalent. $(A \leftrightarrow B) \leftrightarrow C$ is the same as $((A \to B) \land (B \to A)) \to 
          % ^ that was supposed to be a thing with equivalences but I'm too tired to make it work
          $(A \to B) \to C$ and $A \to (B \to C)$ are not logically equivalent because $(F \to T) \to F$ is false while $F \to (T \to F)$ is true.
    \item Suppose C is a tautology. What can you say about the argument \begin{tabular}{c}A\\B\\\hline C\end{tabular}? (M 46)\\
          The argument is valid. Because C is a tautology, the conclusion of the argument is always true, and thus there can be no counterexample to the argument.
    \item Suppose that A and B are logically equivalent. What can you say about $A \lor B$? (M 46)\\
          A and B are each logically equivalent to $A \lor B$. If A is true, $A \lor B$ is also true. If A is false, B must be false as well (since $A \leftrightarrow B$), and thus $A \lor B$ is also false. This means A is logically equivalent to $A \lor B$, and because B is logically equivalent to A, B is logically equivalent to $A \lor B$ as well.
    \item Suppose that A and B are not logically equivalent. What can you say about $A \lor B$? (M 46)\\
          $A \lor B$ is a tautology. Where A is true, B must be false and $A \lor B$ must be true. Where A is false, B must be true and so $A \lor B$ must be true. $A \lor \neg A$ is a tautology, and because $A \lor B$ follows from it, $A \lor B$ must also be a tautology.
\end{enumerate}
\end{document}

### 2023-09-06 Submission 1.2.tex
Submission 1.2 - it's a lot of HWs together, I was sick, I don't know what's going on

@@
 
+\begin{enumerate}
+    \item Jones, feeling upset about the insecurity of the Social Security system, sighs that he faces a dilemma: “If taxes aren't raised, I'll have no money when I'm old.
+If taxes are raised, I'll have no money now.” Smith, ever the even-tempered one, reasons that neither of Jones's contentions is true.
+Jones answers, “Aha!
+You've contradicted yourself!” Show that Smith's assertion that both Jones's claims are false is indeed contradictory.
+(B 133)\\
+          A: Taxes are raised.\\
+          B: Jones will have no money when he's old.\\
+          C: Jones will have no money now.\\
+          D: Both of Jones' claims are false, i.e. $\neg (A \to C) \land \neg (\neg A \to B)$\\
+          \begin{tabular}{c|c|c|c|c|c}
+              A & B & C & $A \to C$ & $\neg A \to B$ & D \\
+              \hline
+              T & T & T & T         & T              & F \\
+              T & T & F & F         & T              & F \\
+              T & F & T & T         & T              & F \\
+              T & F & F & F         & T              & F \\
+              F & T & T & T         & T              & F \\
+              F & T & F & T         & T              & F \\
+              F & F & T & T         & F              & F \\
+              F & F & F & T         & F              & F \\
+          \end{tabular}\\
+          There is no case in which both of Jones' claims are false, so Smith's assertion is never true. Thus, Smith's assertion is a contradiction.
+    \item Consider the statement: If a fetus is a person, it has a right to life. Which of the following sentences follow from this? (B 26)\\
+          Statement C, "If a fetus doesn't have a right to life, it isn't a person" and Statement E, "A fetus isn’t a person only if it doesn’t have a right to life," follow by contraposition.
+    \item Consider this statement from IRS publication 17: Your Federal Income Tax: If you are single, you must file a return if you had gross income of \$3,560 or more for the year.
+What follows from this, together with the information listed?
+(B 26)
+          \begin{enumerate}
+              \item You are single with an income of \$2,500.\\
+                    No information can be derived from this.
+              \item You are married with an income of \$2,500.\\
+                    No information can be derived from this.
+              \item You are single with an income of \$25,000.\\
+                    You must file a tax return.
+              \item You are single but do not have to file a return.\\
+                    Your gross income was less than \$3,560.
+              \item You are married but do not have to file a return.\\
+                    No information can be derived from this.
+          \end{enumerate}
+\end{enumerate}
+
+\begin{flushleft}
+    For problems 4-9, prove that the argument is valid (if applicable) with a proof by natural deduction.
+\end{flushleft}
+\begin{enumerate}
+    \setcounter{enumi}{3}
+    \item I have already said that he must have gone to King's Pyland or to Mapleton. He is not at King's Pyland, therefore he is at Mapleton. (Sir Arthur Conan Doyle) (B 20)
+          \begin{enumerate}
+              \item He is at King's Pyland or at Mapleton. (Premise 1)
+              \item He is not at King's Pyland. (Premise 2)
+              \item He is at Mapleton. (DS)
+          \end{enumerate}
+    \item If I'm right, then I'm a fool. But if I'm a fool, I'm not right. Therefore, I'm not right. (B 70)
+          \begin{enumerate}
+              \item If I'm right, then I'm a fool. (Premise 1)
+              \item If I'm a fool, then I'm not right. (Premise 2)
+              \item If I'm not a fool, then I'm not right. (ContraPos, b)
+              \item I'm either a fool or not a fool. (Taut)
+              \item I'm not right or I'm not right. (CD, b, c, d)
+              \item I'm not right. (Rep, e)
+          \end{enumerate}
+    \item Congress will agree to the cut only if the President announces his support first. The President won't announce his support first, so Congress won't agree to the cut. (B 20)
+          \begin{enumerate}
+              \item Congress will agree to the cut only if the President announces his support first. (Premise 1)
+              \item The President will not announce his support first. (Premise 2)
+              \item If the president does not announce his support, Congress will not agree to the cut. (ContraPos, a)
+              \item Congress will not agree to the cut. (MP, b, c)
+          \end{enumerate}
+    \item If you are ambitious, you'll never achieve all your goals. But life has meaning only if you have ambition. Thus, if you achieve all your goals, life has no meaning. (B 132)
+          \begin{enumerate}
+              \item If you are ambitious, then you'll never achieve all your goals. (Premise 1)
+              \item Life has meaning only if you have ambition. (Premise 2)
+              \item If you achieve all your goals, you are not ambitious. (ContraPos, a)
+              \item If you are not ambitious, life does not have meaning. (ContraPos, b)
+              \item If you achieve all your goals, life does not have meaning. (HS, c, d)
+          \end{enumerate}
+    \item Mittens meows exactly when she is hungry. Mittens is meowing, but she isn't hungry. Therefore the end of the Earth is at hand. (B 70)
+          \begin{enumerate}
+              \item Mittens meows exactly when she is hungry. (Premise 1)
+              \item Mittens is meowing, but she isn't hungry. (Premise 2)
+              \item The world is ending. (ContraPrm, a, b)
+          \end{enumerate}
+    \item God is omnipotent if and only if He can do everything.
+If He can't make a stone so heavy that He can't lift it, then he can't do everything.
+But if He can make a stone so heavy that He can't lift it, He can't do everything.
+Therefore, either God is not omnipotent or God does not exist.
+(B 132)
+          \begin{enumerate}
+              \item God is omnipotent if and only if he can do everything. (Premise 1)
+              \item If he can't make a stone so heavy that he can't lift it, then he can't do everything. (Premise 2)
+              \item If he can make a stone so heavy that he can't lift it, then he can't do everything. (Premise 3)
+              \item He either can or can't make a stone so heavy that he can't lift it. (Taut)
+              \item He can't do everything or he can't do everything. (CD, b, c, d)
+              \item He can't do everything. (Rep, e)
+              \item If God is omnipotent, he can do everything, and if he can do everything, he is omnipotent. (Equiv, a)
+              \item If God is omnipotent, he can do everything. (Simp, g)
+              \item If he can't do everything, he is not omnipotent. (ContraPos, h)
+              \item He is not omnipotent. (MP, i, f)
+              \item He is not omnipotent or he does not exist. (Add, j)
+          \end{enumerate}
+\end{enumerate}
+
 \begin{flushleft}
@@
 \begin{itemize}
-    \item \textit{If the claim that’s made is correct, prove that it’s correct.}
-    \item \textit{If the claim that’s made is incorrect, prove that it’s incorrect.}
+    \item \textit{If the claim that's made is correct, prove that it's correct.}
+    \item \textit{If the claim that's made is incorrect, prove that it's incorrect.}
     \item \textit{If the problem asks a question, answer it with a correct claim, and prove that your claim is correct.}
@@
 $A \lor \neg A$ is a tautology, and because $A \lor B$ follows from it, $A \lor B$ must also be a tautology.
+    \item There are a number of languages with only two operators that are equivalent to truth-functional logic.
+Show that it is sufficient to have only the negation (¬) and the conditional (→) by writing sentences (containing only the operators ¬ and →) that are logically equivalent to the following: (M 46-7)
+          \begin{itemize}
+              \item $A \lor B$
+              \item $A \land B$
+              \item $A \leftrightarrow B$
+          \end{itemize}
+          $A \lor B$ is true where A is true or where B is true. $A \to B$ is true where A is false or where B is true. Therefore, $A \lor B$ is equivalent to $\neg A \to B$.\\
+          $A \land B$ can be found using DeMorgan's Law.
+$A \land B$ is equivalent to $\neg \neg(A \land B)$, which is in turn equivalent to $\neg(\neg A \lor \neg B)$.
+This can be rewritten using the definition for $\lor$ above to get $\neg (A \to \neg B)$.\\
+          $A \leftrightarrow B$ is defined as $(A \to B) \land (B \to A)$. Using the definition above, this can be rewritten as $\neg ((A \to B) \to \neg (B \to A))$.
+    \item Show that there is a language containing only two truth-functional operators, the negation ($\neg$) and the disjunction ($\lor$), that is equivalent to truth-functional logic. (M 47)\\
+          By the same logic as above, $A \to B$ can be written as $\neg A \lor B$.\\
+          Again using DeMorgan's Law, $A \land B$ can be written as $\neg(\neg A \lor \neg B)$.\\
+          $A \leftrightarrow B$ is defined as $(A \to B) \land (B \to A)$, which can be rewritten as $\neg (\neg (\neg A \lor B) \lor \neg (\neg B \lor A))$.
 \end{enumerate}

### 2023-09-16 Submission 2.1.tex
Day 12 HW

\documentclass{article}
\usepackage{enumerate}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage[table]{xcolor}
\usepackage{graphicx}
\graphicspath{ {./images/} }

\title{Submission 2.1}
% \author{Saffron Liu}
\date{}
\begin{document}
\maketitle

\begin{enumerate}
    \item Translate into smooth English: $\forall x\forall y((Px \land Ty \land Dxy \land Oxy) \to \neg \exists z(Pz \land Kzxy))$.\\
          Let “Px” mean “x is a person”, “Tx” mean “x is a time”, “Dxy” mean “x is down at time y”, “Oxy” mean “x is out at time y”, and “Kxyz” mean “x knows y at time z”.\\
          Every pair of individuals x and y is such that if x is a person, y is a time, x is down at time y, and x is out at time y, then there is no z such that if z is a person, z knows x at time y. More English-y: If someone is down and out, they cannot be known by someone else at that time. Even more English-y: Nobody knows you when you're down and out.
\end{enumerate}

\begin{flushleft}
    For problems 2-15: \textit{Write a sentence in quantificational logic that captures as much of the given information as possible. Remember to delineate the extensions you assign to names and predicates.}
\end{flushleft}
\begin{enumerate}
    \setcounter{enumi}{1}
    \item All's well that ends well. (Shakespeare) (B 148)\\
          Ax: x is well\\
          Bx: x ends well\\
          $\forall x(Bx \to Ax)$
    \item The things which are seen are temporal; but the things which are not seen are eternal. (II Corinthians 4:18) (B 169)\\
          Ax: x is seen\\
          Bx: x is eternal\\
          $\forall x((Ax \to \neg Bx) \land (\neg Ax \to Bx))$, which is equivalent to $\forall x (Ax \leftrightarrow \neg Bx)$.
    \item \includegraphics{2.1.4} \\
          Ax: x is in A\\
          Bx: x is in B\\
          $\forall x(Ax \to Bx)$
    \item \includegraphics{2.1.5}\\
          Ax: x is in A\\
          Bx: x is in B\\
          $\forall x(Ax \leftrightarrow \neg Bx)$
    \item If you don't love yourself, you can't love anybody else.\\
          Lxy: x loves y\\
          y: you\\
          $\neg Lyy \to \neg \exists x(Lyx)$
    \item NSYNC is the best band ever.\\
          n: NSYNC\\
          A: is the best band\\
          $An$
    \item Somebody loves everybody.\\
          Axy: x loves y\\
          $\exists x \forall y (Axy)$
    \item There is someone for everybody.\\
          Axy: There is x for y\\
          $\exists x \forall y(Axy)$
    \item Scrooge doesn't love anybody.\\
          Axy: x loves y\\
          s: Scrooge\\
          $\neg \exists x (Lsx)$
    \item Only the shallow know themselves. (Oscar Wilde) (B 169)\\
          Ax: x is shallow\\
          Bx: x knows itself\\
          $\forall x(Bx \to Ax)$
    \item Everybody has a mother.\\
          Ax: x has a mother\\
          $\forall x(Ax)$
\end{enumerate}

\end{document}

### 2023-09-19 Submission 2.1.tex
Fix question 9
Day 13 HW

@@
             Axy: There is x for y\\
-          $\exists x \forall y(Axy)$
+            $\forall y \exists x(Axy)$\\
+            \textit{Note: the order of quantifiers matters!}
       \item Scrooge doesn't love anybody.\\
@@
             $\forall x(Ax)$
+      \item There are at least two pigs.\\
+            Px: x is a pig\\
+            $\exists x \exists y (Px \land Py \land (x \neq y))$
+      \item There are exactly two pigs.\\
+            $\exists x \exists y \exists z (Px \land Py \land x \neq y \land ((z \neq x \land z \neq y) \to \neg Pz))$
+      \item There are at most two pigs.\\ 
+            $\neg \exists x \exists y \exists z (Px \land Py \land Pz \land (x \neq y \land x \neq z \land y \neq z))$
+            \textit{(there are not 3 unique pigs)}
+
 \end{enumerate}

### 2023-09-27 Submission 2.2.tex
Day 15 HW

\documentclass{article}
\usepackage{enumerate}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage[table]{xcolor}

\title{Submission 2.1}
% \author{Saffron Liu}
\date{}
\begin{document}
\maketitle

\begin{flushleft}
      For problems 4-12:
\end{flushleft}
\begin{enumerate}[(a)]
      \item \textit{Translate the argument into quantificational logic. Be sure to delineate the extensions you give to predicates and names, and write the argument (now in quantificational logic) in standard form.}
      \item \textit{Claim whether the argument is valid or sound. In some cases, soundness will be difficult to determine, so “soundness is difficult to determine” is an appropriate answer.}
      \item \textit{Prove that the argument is valid or invalid (as appropriate) with an informal proof.}
      \item \textit{Prove that the argument is valid (if applicable) with a proof by natural deduction.}
\end{enumerate}

\begin{enumerate}
      \setcounter{enumi}{3}
      \item Everything has a cause. Therefore something is the cause of everything. (Some people think St. Thomas Aquinas advocated this.) (B 183)
            \begin{enumerate}
                  \item Cxy: x is the cause of y\\

                        $\forall x \exists y(Cyx)$\\
                        \rule{15em}{.5pt}\\
                        $\exists y \forall x(Cyx)$
                  \item This argument is invalid, and therefore unsound.
                        % Actually this one might be valid?
                  \item Although each thing may have a cause, they may not be the same cause. Therefore, it does not follow that all things share the same cause. For example, if A is caused by A, B is caused by A, C is caused by D, and D is caused by A, the premise is satisfied, but the conclusion is not.
            \end{enumerate}
      \item Fred hates everyone who hates Al. Al hates everyone. So Al and Fred hate each other. (B213)
            \begin{enumerate}
                  \item Hxy: x hates y\\
                        a: Al\\
                        f: Fred\\

                        $\forall x(Hax)$\\
                        $\forall y(Hya \to Hfy)$\\
                        \rule{15em}{.5pt}\\
                        $Haf \land Hfa$
                  \item This argument is valid. Soundness is difficult to determine.
            \end{enumerate}
      \item All insects in this house are large and hostile. Some insects in this house are impervious to pesticides. Thus, some large, hostile insects are impervious to pesticides. (B 213)
            \begin{enumerate}
                  \item Lx: x is large\\
                        Hx: x is hostile\\
                        Px: x is impervious to pesticides\\

                        $\forall x(Lx \land Hx)$\\
                        $\exists x(Px)$\\
                        \rule{15em}{.5pt}\\
                        $\exists x(Lx \land Hx \land Px)$
                  \item This argument is valid and (unfortunately) quite likely sound.
            \end{enumerate}
      \item Some students cannot succeed at the university. All students who are bright and mature can succeed. It follows that some students are either not bright or immature. (B 213)
            \begin{enumerate}
                  \item Sx: x can succeed\\
                        Bx: x is bright\\
                        Mx: x is mature\\

                        $\exists x(\neg Sx)$\\
                        $\forall x((Bx \land Mx) \to Sx)$\\
                        \rule{15em}{.5pt}\\
                        $\exists x(\neg Bx \lor \neg Mx)$
                  \item This argument is valid. I would argue that it's unsound.
            \end{enumerate}
      \item There are at least 3 pigs. So there are at least two pigs.
            \begin{enumerate}
                  \item Px: x is a pig\\
                        $\exists x \exists y \exists z (Px \land Py \land Pz \land (x \neq y) \land (x \neq z) \land (y \neq z))$\\
                        \rule{15em}{.5pt}\\
                        $\exists x \exists y (Px \land Py \land (x \neq y))$
                  \item This argument is valid and sound.
            \end{enumerate}
      \item Popeye and Olive Oyl like each other since Popeye likes everyone who likes Olive Oyl, and Olive Oyl likes everyone. (B 218)
            \begin{enumerate}
                  \item Lxy: x likes y\\
                        p: Popeye\\
                        o: Olive Oyl\\

                        $\forall x(Lxo \to Lpx)$\\
                        $\forall x(Lox)$\\
                        \rule{15em}{.5pt}\\
                        $Lpo \land Lop$
                  \item This argument is valid and sound.
            \end{enumerate}
      \item This argument is unsound, for its conclusion is false, and no sound argument has a false conclusion. (J 49)
            \begin{enumerate}
                  \item U: x is unsound\\
                        F: x has a false conclusion\\
                        a: this argument\\

                        $Fa$\\
                        $\forall x(Fx \to Ux)$\\
                        \rule{15em}{.5pt}\\
                        $Ua$
                  \item This argument is valid. Soundness is (very) difficult to determine.
                  \item The second premise states that any argument with a false conclusion is unsound. The first premise states that "this argument" has a false conclusion, so it must follow that "this argument" is unsound.
            \end{enumerate}
      \item Everyone likes Mandy. Mandy likes nobody but Andy. Therefore Mandy and Andy are the same person. (B 238)
            \begin{enumerate}
                  \item Lxy: x likes y\\
                        m: Mandy\\
                        a: Andy\\

                        $\forall x(Lxm)$\\
                        $\forall x(x \neq a \to \neg Lmx)$\\
                        \rule{15em}{.5pt}\\
                        $m=a$
                  \item This argument is valid and sound.
                        % Everyone (including Mandy) likes Mandy; Mandy likes nobody except Andy. Mandy must like Mandy, so Andy must be Mandy.
            \end{enumerate}
      \item Everyone is afraid of Mr. Hyde. Mr. Hyde is afraid only of Dr. Jekyll. Therefore, Dr. Jekyll is Mr. Hyde. (B 234)
            \begin{enumerate}
                  \item Axy: x is afraid of y\\
                        h: Mr. Hyde\\
                        j: Dr. Jekyll\\

                        $\forall x(Axh)$\\
                        $\forall x (x \neq j \to \neg Ahx)$\\
                        \rule{15em}{.5pt}\\
                        $h=j$
                  \item This argument is valid and sound.
            \end{enumerate}
\end{enumerate}

\begin{flushleft}
      For problems 13-15:
\end{flushleft}
\begin{enumerate}[(a)]
      \item \textit{Claim whether the argument is valid or invalid.}
      \item \textit{Prove that the argument is valid or invalid (as appropriate) with an informal proof.}
\end{enumerate}

\begin{enumerate}
      \setcounter{enumi}{12}
      \item \begin{tabular}{c}
                  $\forall x(Fx \to Gx)$ \\
                  $\forall x(Fx \to Hx)$ \\
                  \hline
                  $\forall x(Gx \to Hx)$
            \end{tabular}
            (B 204)

            \begin{enumerate}
                  \item The argument is invalid.
                  \item Let y be an instance of x such that Fy is false, Gy is true, and Hy is false. In this case, both $Fy \to Gy$ and $Fy \to Hx$ are true, but $Gy \to Hy$ is not. This means that the conclusion is not true for all x.
            \end{enumerate}
      \item \begin{tabular}{c}
                  $\forall x(Fx \to Gx)$ \\
                  \hline
                  $\forall x(\neg Gx \to \neg Fx)$
            \end{tabular}
            (B 204)
            \begin{enumerate}
                  \item This argument is valid.
                  \item This is a contrapositive.
            \end{enumerate}
      \item \begin{tabular}{c}
                  $\neg \exists x(Fx \land Gx)$ \\
                  $\forall x(Gx \to Hx)$        \\
                  \hline
                  $\forall x(Fx \to \neg Hx)$
            \end{tabular}
            (B 204)
            \begin{enumerate}
                  \item This argument is invalid.
                  \item Let y be an instance of x such that Fy is true and Hy is true. Since Fy is true and there is no x such that both Fx and Gx are true, Gy must be false. However, this does not imply anything abuot Hy. Since this case does not pose any contradictions but renders the conclusion false, the argument is invalid.
            \end{enumerate}

\end{enumerate}

\end{document}

### 2023-09-27 Submission 2.2.tex
Day 16 HW

@@
 \maketitle
+
+\begin{enumerate}
+      \item In school, a curve is a function, usually from [0, 100] to [0, 100]. The following are some made-up definitions:
+            \begin{itemize}
+                  \item A curve f is called fair if $\forall x \forall y (x \geq y \to f(x) \geq f(y))$.
+                  \item A curve f is called totally unfair if $\forall x \forall y (x > y \to f(x) < f(y))$.
+                  \item A curve f is called progressive if $\forall x \forall y (x < y \to f(x) - x > f(y) - y)$.
+            \end{itemize}
+            \begin{enumerate}
+                  \item Say (both in English and in Quantificational Logic) what it means for a curve to be unfair.\\
+                        A curve is unfair if a grade x greater than or equal to a grade y before the curve ends up lower than grade y after the curve.
+$\exists x \exists y \neg(x \geq y \to f(x) \geq f(y))$
+                  \item Say (both in English and in Quantificational Logic) what it means for a curve to be not totally unfair.\\
+                        A curve is not totally unfair if a score x greater than score y does not end up lower after the curve. $\exists x \exists y \neg(x > y \to f(x) < f(y))$
+                  \item Conjure an example of a curve that is unfair but not totally unfair.\\
+                        An example of this would be f(x) = (x \% 2) * 100.
+This is not fair, because when a = 2 and b = 1 (i.e.
+$a \geq b$), f(a) = 0 and f(b) = 100 (i.e.
+$f(a) < f(b)$).
+However, this curve is also not totally unfair, because when a = 3 and b = 2 (i.e.
+$a > b$), f(a) = 100 and f(b) = 0 (i.e.
+$f(a) \geq f(b)$).
+                  \item Say in English what it means for a curve to be progressive. Say (both in English and in Quantificational Logic) what it means for a curve to be not progressive.\\
+                        A curve is progressive if the difference between the post- and pre- curve scores is higher for lower scores (for most intents and purposes, if lower scores are "bumped" more than higher scores.)\\
+                        A curve is not progressive if there is a case where a lower score is bumped the same amount or less than a higher score.
+$\exists x \exists y (x < y \to f(x) - x \leq f(y) - y)$
+                  \item Classify the curve “everyone gets 5 points” as fair or unfair and progressive or not progressive.\\
+                        This curve is fair but not progressive.
+            \end{enumerate}
+      \item The limit of a function, f(x), at a point c is l if and only if for all $\varepsilon > 0$ there is a $\delta > 0$
+            such that, for all x, if the distance between x and c is less than $\delta$, then the distance between
+            f(x) and l is less than $\varepsilon$. Suppose f is a function, and the limit of f at 0 is not 1. Given
+            this definition of a limit, how can you prove that the limit of f at 0 is not 1? That is, what
+            evidence would you need to provide in order to prove that the limit of f at 0 is not 1?\\
+            You would need to find an $\varepsilon$ such that for all $\delta$, the distance between x and 1 being less than $\delta$ does not guarantee that the distance between f(x) and 1 will be less than $\varepsilon$.
+      \item A function, f, is continuous at a point c if and only if for all sequences of real numbers,
+      $x_{n}$, that converge to c, the sequence $f(x_{n})$ converges to f(c). Suppose f is a function, and
+      f (0) = 1. Given this definition of pointwise continuity above, what would you need to do
+      to prove that f is discontinuous at 0? That is, what evidence would you need to provide in
+      order to prove that f is discontinuous at 0?\\
+      You would need to find a sequence of real numbers $x_{n}$ that converges to 0 but where $f(x_{n})$ does not converge to 1.
+\end{enumerate}
 

### 2023-10-02 Submission 2.2.tex
Day 17 HW

@@
                   \item This argument is invalid, and therefore unsound.
-                        % Actually this one might be valid?
                   \item Although each thing may have a cause, they may not be the same cause.
 Therefore, it does not follow that all things share the same cause.
-For example, if A is caused by A, B is caused by A, C is caused by D, and D is caused by A, the premise is satisfied, but the conclusion is not.
+For example, if A is caused by A, and B is caused by B, there is not a single thing that causes both A and B.
+                  \item Because the argument is invalid, it cannot be proven with natural deduction.
             \end{enumerate}
@@
                   \item This argument is valid. Soundness is difficult to determine.
+                        \setcounter{enumii}{3}
+                  \item \begin{enumerate}
+                              \item $\forall x(Hax)$ (Premise 1)
+                              \item $\forall y(Hya \to Hfy)$ (Premise 2)
+                              \item $Haf$ (UI, i)
+                              \item $Haa$ (UI, i)
+                              \item $Haa \to Hfa$ (UI, ii)
+                              \item $Hfa$ (MP, v, iv)
+                              \item $Haf \land Hfa$ (Conj, iii, vi)
+                        \end{enumerate}
             \end{enumerate}
@@
                   \item This argument is valid and (unfortunately) quite likely sound.
+                        \setcounter{enumii}{3}
+                  \item \begin{enumerate}
+                              \item $\forall x(Lx \land Hx)$ (Premise 1)
+                              \item $\exists x(Px)$ (Premise 2)
+                              \item $Pa$ (EI, ii)
+                              \item $La \land Ha$ (UI, i)
+                              \item $La \land Ha \land Pa$ (Conj, iii, iv)
+                              \item $\exists x(Lx \land Hx \land Px)$ (EG, v)
+                        \end{enumerate}
             \end{enumerate}
@@
                   \item This argument is valid. I would argue that it's unsound.
+                        \setcounter{enumii}{3}
+                  \item \begin{enumerate}
+                              \item $\exists x(\neg Sx)$ (Premise 1)
+                              \item $\forall x((Bx \land Mx) \to Sx)$ (Premise 2)
+                              \item $\neg Sa$ (EI, i)
+                              \item $(Ba \land Ma) \to Sa$ (UI, ii)
+                              \item $\neg Sa \to \neg (Ba \land Ma)$ (ContraPos, iv)
+                              \item $\neg Sa \to (\neg Ba \lor \neg Ma)$ (DeM, v)
+                              \item $\neg Ba \lor \neg Ma$ (MP, iii, vi)
+                              \item $\exists x(\neg Bx \lor \neg Mx)$ (EG, vii)
+                        \end{enumerate}
             \end{enumerate}
@@
                   \item This argument is valid and sound.
+                        \setcounter{enumii}{3}
+                  \item \begin{enumerate}
+                              \item $\exists x \exists y \exists z (Px \land Py \land Pz \land (x \neq y) \land (x \neq z) \land (y \neq z))$ (Premise 1)
+                              \item $\exists y \exists z (Pa \land Py \land Pz \land (a \neq y) \land (a \neq z) \land (y \neq z))$ (EI, i)
+                              \item $\exists z (Pa \land Pb \land Pz \land (a \neq b) \land (a \neq z) \land (b \neq z))$ (EI, ii)
+                              \item $Pa \land Pb \land Pc \land (a \neq b) \land (a \neq c) \land (b \neq c)$ (EI, iii)
+                              \item $Pa \land Pb \land Pc \land (a \neq b) \land (a \neq c)$ (Simp, iv)
+                              \item $Pa \land Pb \land Pc \land (a \neq b)$ (Simp, v)
+                              \item $Pa \land Pb \land (a \neq b)$ (Simp, vi)
+                              \item $\exists y(Pa \land Py \land (a \neq y))$ (EG, vii)
+                              \item $\exists x \exists y(Px \land Py \land (x \neq y))$ (EG, viii)
+                        \end{enumerate}
             \end{enumerate}
@@
                   \item This argument is valid and sound.
+                        \setcounter{enumii}{3}
+                  \item \begin{enumerate}
+                              \item $\forall x(Lxo \to Lpx)$ (Premise 1)
+                              \item $\forall x(Lox)$ (Premise 2)
+                              \item $Loo \to Lpo$ (UI, i)
+                              \item $Loo$ (UI, ii)
+                              \item $Lpo$ (MP, iii, iv)
+                              \item $Lop$ (UI, ii)
+                              \item $Lpo \land Lop$ (Conj, v, vi)
+                        \end{enumerate}
             \end{enumerate}
@@
 The first premise states that "this argument" has a false conclusion, so it must follow that "this argument" is unsound.
+                  \item \begin{enumerate}
+                              \item $Fa$ (Premise 1)
+                              \item $\forall x(Fx \to Ux)$ (Premise 2)
+                              \item $Fa \to Ua$ (UI, ii)
+                              \item $Ua$ (MP, iii, i)
+                        \end{enumerate}
             \end{enumerate}
@@
                   \item This argument is valid and sound.
+                        \setcounter{enumii}{3}
+                  \item \begin{enumerate}
+                              \item $\forall x(Lxm)$ (Premise 1)
+                              \item $\forall x(x \neq a \to \neg Lmx)$ (Premise 2)
+                              \item $Lmm$ (UI, i)
+                              \item $m \neq a \to \neg Lmm$ (UI, ii)
+                              \item $Lmm \to m = a$ (ContraPost, iv)
+                              \item $m = a$ (MP, v, iii)
+                        \end{enumerate}
                         % Everyone (including Mandy) likes Mandy; Mandy likes nobody except Andy. Mandy must like Mandy, so Andy must be Mandy.
@@
                   \item This argument is valid and sound.
+                        \setcounter{enumii}{3}
+                  \item \begin{enumerate}
+                              \item $\forall x(Axh)$ (Premise 1)
+                              \item $\forall x (x \neq j \to \neg Ahx)$ (Premise 2)
+                              \item $Ahh$ (UI, i)
+                              \item $h \neq j \to \neg Ahh$ (UI, ii)
+                              \item $Ahh \to h = j$ (ContraPos, iv)
+                              \item $h = j$ (MP, v, iii)
+                        \end{enumerate}
             \end{enumerate}

### 2023-10-25 Munkres 1.1.tex
Munkres 1.1

\documentclass{article}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage{amssymb}

\title{Submission 3.1}
\date{}
\begin{document}
\maketitle

\begin{enumerate}
      \item Check the distributive laws for $\cup$ and $\cap$ and DeMorgan's Laws.
            \begin{enumerate}
                  \item $A \cap (B \cup C) = (A \cap B) \cup (A \cap C)$\\
                        Let $x \in A \cap (B \cup C)$. It can then be said that $x \in A \land (x \in B \lor x \in C)$. Therefore, $(x \in A \land x \in B) \lor (x \in A \land x \in C)$. So, $(x \in A \cap B) \lor (x \in A \cap C)$, so $x \in (A \cap B) \cup (A \cap C)$. It can thus be said that $A \cap (B \cup C) \subset (A \cap B) \cup (A \cap C)$, and because all the steps are reversible, $A \cap (B \cup C) \supset (A \cap B) \cup (A \cap C)$. Therefore, $A \cap (B \cup C) = (A \cap B) \cup (A \cap C)$. (The same reasoning applies for (b) - (c) as well.)
                  \item $A \cup (B \cap C) = (A \cup B) \cap (A \cup C)$\\
                        Let $x \in A \cup (B \cap C)$. It can then be said that $x \in A \lor (x \in B \land x \in C)$. Therefore, $(x \in A \lor x \in B) \land (x \in A \lor x \in B)$. So, $x \in (A \cup B) \lor x \in (A \cup C)$, so $x \in (A \cup B) \cap (A \cup C)$.
                  \item $A - (B \cup C) = (A - B) \cap (A - C)$\\
                        Let $x \in (A - (B \cup C))$. It can then be said that $x \in A \land x \notin B \land x \notin C$. Thus, $(x \in A \land x \notin B) \land (x \in A \land x \notin C)$. Therefore, $x \in A - B \land x \in A - C$, so $x \in (A - B) \cap (A - C)$.
                  \item $A - (B \cap C) = (A - B) \cup (A - C)$\\
                        Let $x \in A - (B \cap C)$. It can then be said that $x \in A \land (x \notin B \lor x \notin C)$. Therefore, $(x \in A \land x \notin B) \lor (x \in A \land x \notin B)$. So $x \in (A - B) \lor x \in (A - C)$, so $x \in (A - B) \cup (A - C)$.
            \end{enumerate}
      \item Determine which of the following statements are true for all sets A, B, C, and D. If a double implication fails, determine whether one or the other of the possible implications holds. If an equality fails, determine whether the statement becomes true if the “equals” symbol is replaced by one or the other of the inclusions symbols $\subset$ or $\supset$.
            \begin{enumerate}
                  \item $(A \subset B$ and $A \subset C) \Leftrightarrow A \subset (B \cup C)$ is not true. However, $(A \subset B$ and $A \subset C) \Rightarrow A \subset (B \cup C)$ is.\\
                  $A \subset B$ and $B \subset B \cup C$. Because $A$ is a subset of something fully contained in $B \cup C$, it must also be a subset of $B \cup C$, so the forward implication works. Let $A = \{0, 1\}$, $B = \{0\}$, and $C = \{1\}$. Here, $A \subset B \cup C$, but $A \not\subset B$ and $A \not\subset C$, so the backwards implication fails.
                        \setcounter{enumii}{4}
                  \item $A - (A - B) = B$ is not true, but $A - (A - B) \subset B$ is.\\
                  $A - B$ = $\{x \mid x \in A \land x \notin B\}$, so $A - (A - B)$ = $\{x \mid x \in A \land \neg (x \in A \land x \notin B)\}$, so $\{x \mid x \in A \land (x \notin A \lor x \in B)\}$. This means $\{x \mid (x \in A \land x \notin A) \lor (x \in A \land x \in B)\}$. $(x \in A \land x \notin A)$ is a contradiction, so the statement simplifies to $\{x \mid x \in A \land x \in B\}$. This means everything within $A - (A - B)$ is contained within $B$, so $A - (A - B) \subset B$ is true, but if some element is contained within $B$ but not $A$, it will not be contained in $A - (A - B)$, so the equality fails.
                        \setcounter{enumii}{8}
                  \item $(A \cap B) \cup (A - B) = A$ is true.\\
                  $(A \cap B) \cup (A - B) = \{x \mid (x \in A \land x \in B) \lor (x \in A \land x \notin B)\}$. Therefore, $\{x \mid x \in A \land (x \in B \lor x \notin B)\}$. ($x \in B \lor x \notin B)$ is a tautology, so the statement simplifies to $\{x \mid x \in A\}$, which is the definition for the set $A$. All steps are reversible, so the equality holds.
                        \setcounter{enumii}{14}
                  \item $A \times (B-C) = (A \times B) - (A \times C)$ is true.\\
                  $A \times (B-C) = \{(a, b) \mid a \in A \land (b \in B \land b \notin C)\}$\\
                  $(A \times B) - (A \times C) = \{(a, b) \mid (a \in A \land b \in B) \land \neg (a \in A \land b \in C)\} = \{(a, b) \mid (a \in A \land b \in B) \land (a \notin A \lor b \notin C)\} = \{(a, b) \mid (a \in A \land b \in B \land a \notin A) \lor (a \in A \land b \in B \land b \notin C)\}$. The first part of the main or contains a contradiction, so the statement simplifies to $\{(a, b) \mid a \in A \land b \in B \land b \notin C\}$. All steps are reversible (I really hope), so the equality holds.
            \end{enumerate}
      \item
            \begin{enumerate}
                  \item Write the contrapositive and converse of the following statement: “If $x < 0$, then $x^{2} - x > 0$,” and determine which (if any) of the three statements are true.\\
                        The statement is true. $x^{2}$ is always positive, and if $x < 0$, then subtracting x will be equivalent to adding a positive number, so $x^{2} - x$ will be greater than 0.\\
                        The contrapositive of the statement is “If $x^{2} - x \leq 0$, then $x \geq 0$" and is also true because it is logically equivalent.\\ 
                        The converse is “If $x^{2} - x > 0$, then $x < 0$,” which is false. If x is 2, $x^{2} - x$ is 2, fulfilling the antecedent but not the consequent.
                  \item Do the same for the statement “If $x > 0$, then $x^{2} - x > 0$.”\\
                        The statement is false. If x is 1, $x^{2} - x$ is 0, so something true implies something false.\\
                        Therefore, the contrapositive, which is “If $x^{2} - x \leq 0$, then $x \leq 0$" is also false.\\
                        The converse, “If $x^{2} - x > 0$, then $x > 0$" is also false. If x is -1, the antecedent is fulfilled but the consequent is not.
            \end{enumerate}
      \item Let A and B be sets of real numbers. Write the negation of each of the following statements:
            \begin{enumerate}
                  \item For every $a \in A$, it is true that $a^{2} \in B$.\\
                        For at least one $a \in A$, it is true that $a^{2} \notin B$.
                  \item For at least one $a \in A$, it is true that $a^{2} \in B$.\\
                        For every $a \in A$, it is true that $a^{2} \notin B$.
                  \item For every $a \in A$, it is true that $a^{2} \notin B$.\\
                        For at least one $a \in A$, it is true that $a^{2} \in B$.
                  \item For every $a \notin A$, it is true that $a^{2} \in B$.\\
                        For at least one $a \notin A$, it is true that $a^{2} \notin B$.
            \end{enumerate}
      \item Let $\mathcal{A}$ be a nonempty collection of sets. Determine the truth of each of the following statements and their converses.
            \begin{enumerate}
                  \item $x \in \bigcup_{A \in \mathcal{A}} A \Rightarrow x \in A$ for at least one $A \in \mathcal{A}$.\\
                        This is true by the definition of an arbitrary union. The converse, “$x \in A$ for at least one $A \in \mathcal{A} \Rightarrow x \in \bigcup_{A \in \mathcal{A}} A$," is also true, again by definition.
                  \item $x \in \bigcup_{A \in \mathcal{A}} A \Rightarrow x \in A$ for every $A \in \mathcal{A}$.\\
                        This is false. A counterexample is given by $x = 0, A = \{0\}, B = \{1\}, \mathcal{A} = \{A, B\}$.\\ 
                        However, the converse, “$x \in A$ for every $A \in \mathcal{A} \Rightarrow x \in \bigcup_{A \in \mathcal{A}} A$," is true. If $x \in A$ for every $A \in \mathcal{A}$, then $x \in A$ for at least one $A \in \mathcal{A}$, which by definition implies that $x \in \bigcup_{A \in \mathcal{A}} A$.
                  \item $x \in \bigcap_{A \in \mathcal{A}} A \Rightarrow x \in A$ for at least one $A \in \mathcal{A}$.\\
                        This is true. By definition, $x \in \bigcap_{A \in \mathcal{A}} A$ implies that $x \in A$ for every $A \in \mathcal{A}$, which in turn implies $x \in A$ for at least one $A \in \mathcal{A}$.\\
                        However, the converse, “$x \in A$ for at least one $A \in \mathcal{A} \Rightarrow x \in \bigcap_{A \in \mathcal{A}} A$," is false. A counterexample is given by $x = 0, A = \{0, 1\}, B = \{1, 2\}, \mathcal{A} = \{A, B\}$. 
                  \item $x \in \bigcap_{A \in \mathcal{A}} A \Rightarrow x \in A$ for every $A \in \mathcal{A}$.\\
                        This is true by the definition of an arbitrary intersection. The converse, “$x \in A$ for every $A \in \mathcal{A} \Rightarrow x \in \bigcap_{A \in \mathcal{A}} A$," is also true by definition.
            \end{enumerate}
      \item Write the contrapositive of each statement in exercise 5.
            \begin{enumerate}
                  \item $x \notin A$ for every $A \in \mathcal{A} \Rightarrow x \notin \bigcup_{A \in \mathcal{A}} A$
                  \item $x \notin A$ for at least one $A \in \mathcal{A} \Rightarrow x \notin \bigcup_{A \in \mathcal{A}} A$
                  \item $x \notin A$ for every $A \in \mathcal{A} \Rightarrow x \notin \bigcap_{A \in \mathcal{A}} A$
                  \item $x \notin A$ for at least one $A \in \mathcal{A} \Rightarrow x \notin \bigcap_{A \in \mathcal{A}} A$
            \end{enumerate}
      \item Given sets A, B, and C, express each of the following sets in terms of A, B, and C, using the symbols $\cup$, $\cap$, and $-$.
            \begin{enumerate}
                  \item $D = \{x \mid x \in A$ and $(x \in B$ or $x \in C)\} = A \cap (B \cup C)$
                  \item $E = \{x \mid (x \in A$ and $x \in B)$ or $x \in C\} = (A \cap B) \cup C$
                  \item $F = \{x \mid x \in A$ and $(x \in B \Rightarrow x \in C)\} = A - (B - C)$\\
                  Let $x \in F$. It can then be said that $x \in A \land (x \in B \to x \in C)$, so $x \in A \land (x \notin B \lor x \in C)$, so $x \in A \land \neg(x \in B \land x \notin C)$. Therefore, $x \in A \land x \notin (B - C)$.
            \end{enumerate}
      \item If a set $A$ has two elements, show that $\mathcal{P}(A)$ has 4 elements. How many elements does $\mathcal{P}(A)$ have if $A$ has one element? Three elements? No elements? Why is $\mathcal{P}(A)$ called the power set of $A$?\\
            Let $A = \{a, b\}$, a set with 2 elements. $\mathcal{P}(A) = \{\varnothing, \{a\}, \{b\}, \{a, b\}\}$, a set with 4 elements.\\
            Let $A = \{a\}$, a set with 1 element. $\mathcal{P}(A) = \{\varnothing, \{a\}\}$, a set with 2 elements.\\
            Let $A = \{a, b, c\}$, a set with 3 elements. $\mathcal{P}(A) = \{\varnothing, \{a\}, \{b\}, \{c\}, \{a, b\}, \{a, c\}, \{b, c\}, \{a, b, c\}\}$, a set with 8 elements.\\
            Let $A = \varnothing$, a set with no elements. $\mathcal{P}(A) = \{\varnothing\}$, a set with 1 element.\\
            $\mathcal{P}(A)$ is called the power set of $A$ because where $A$ has $n$ elements, $\mathcal{P}(A)$ has $2^{n}$ elements. For each set in $\mathcal{P}(A)$, each element of $A$ can be either included or excluded. This results in choosing between one of two options $n$ times, resulting in $2^{n}$ possible options.
      \item Formulate and prove DeMorgan's laws for arbitrary unions and intersections.\\
            $A - \bigcup_{B \in \mathcal{B}} B = \bigcap_{B \in \mathcal{B}}(A - B)$\\
            Let $x \in A - \bigcup_{B \in \mathcal{B}} B$. It can then be said that $x \in A \land \neg (x \in B$ for at least one $B \in \mathcal{B})$. Therefore, $x \in A \land x \notin B$ for every $B \in \mathcal{B}$. So, $x \in (A - B)$ for every $B \in \mathcal{B}$, so $x \in \bigcap_{B \in \mathcal{B}}(A - B)$. Again, all the steps are reversible, showing equality. \\
            $A - \bigcap_{B \in \mathcal{B}} B = A - \bigcup_{B \in \mathcal{B}} B$\\
            Let $x \in A - \bigcap_{B \in \mathcal{B}}$. It can then be said that $x \in A \land \neg(x \in B$ for all $B \in \mathcal{B})$. Therefore, $x \in A \land x \notin B$ for at least one $B \in \mathcal{B}$. So, $x \in A \land x \notin \bigcup_{B \in \mathcal{B}} B$. Therefore, $x \in A - \bigcup_{B \in \mathcal{B}} B$. These steps are reversible, showing equality.
      \item Let $\mathbb{R}$ denote the set of real numbers. For each of the following subsets of $\mathbb{R} \times \mathbb{R}$ determine whether it is equal to the Cartesian product of two subsets of R.
      \begin{enumerate}
            \item \{$(x, y) \mid x$ is an integer\} is the Cartesian product $\mathbb{Z} \times \mathbb{R}$.
            \item \{$(x, y) \mid 0 < y \leq 1$\} is the Cartesian product $\mathbb{R} \times (0, 1]$.
            \item $\{(x, y) \mid y > x\}$ is not a Cartesian product.\\
            (0, 1) and (1, 2) are contained within the given subset. If the subset were a Cartesian product of two subsets of $\mathbb{R}$, (1, 1) and (0, 2) would then also have to be included in the subset. However, (1, 1) is not, so the subset must not be a Cartesian product.
            \item \{$(x, y) \mid $ x is not an integer and y is an integer\} is the Cartesian product $(\mathbb{R} - \mathbb{Z}) \times \mathbb{Z}$.
            \item $\{(x, y) \mid x^{2} + y^{2} < 1\}$ is not a Cartesian product.\\
            (0, 0.75) and (0.75, 0) are both contained within the given subset. If the subset were a Cartesian product of two subsets of $\mathbb{R}$, it would contain (0, 0) and (0.75, 0.75). However, (0.75, 0.75) is not contained, so the subset must not be a Cartesian product.
      \end{enumerate}
\end{enumerate}
\end{document}

### 2023-11-21 Munkres 1.2.tex
Munkres 1.2 - This Should have been 3-5 separate commits but ah... well.

\documentclass{article}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage{amssymb}
\usepackage{graphicx}
\graphicspath{ {./images/} }

\title{Submission 3.2}
\date{}
\begin{document}
\maketitle
\begin{enumerate}
      \item Let $f:A \to B$. Let $A_{0} \subset A$ and $B_{0} \subset B$.
            \begin{enumerate}
                  \item Show that $A_{0} \subset f^{-1}(f(A_{0}))$ and that equality holds if $f$ is injective.\\

                        Proof that $A_{0} \subset f^{-1}(f(A_{0}))$:\\
                        By the definition of an inverse, $f^{-1}(C) = \{x \mid f(x) \in C\}$. Thus, $f^{-1}(f(A_{0})) = \{x \mid f(x) \in f(A_{0})\}$.\\
                        For every $a \in A_{0}$, $f(a) \in f(A_{0})$, fulfilling the conditions of the set.\\
                        Therefore, $A_{0} \subset f^{-1}(f(A_{0}))$.\\

                        Proof that $A_{0} = f^{-1}(f(A_{0}))$ if $f$ is injective:\\
                        By the definition of an inverse, $f^{-1}(f(A_{0})) = \{x \mid f(x) \in f(A_{0})\}$.\\
                        Let $a \in f^{-1}(f(A_{0}))$. Therefore, $f(a) \in f(A_{0})$. Therefore, there is some $x \in A_{0}$ such that $f(a) = f(x)$.\\
                        By the definition of injectivity, $f(a) = f(x)$ implies that $a = x$.\\
                        Therefore, for every $a \in f^{-1}(f(A_{0}))$, $a = x \in A_{0}$. Thus, $f^{-1}(f(A_{0})) \subset A_{0}$. As proven above, $A_{0} \subset f^{-1}(f(A_{0}))$ as well, so $A_{0} = f^{-1}(f(A_{0}))$.

                  \item Show that $f(f^{-1}(B_{0})) \subset B_{0}$ and that equality holds if $f$ is surjective.\\

                        Proof that $f(f^{-1}(B_{0})) \subset B_{0}$:\\
                        By the definition of a function, $f(f^{-1}(B_{0})) = \{x \mid f^{-1}(x) \in f^{-1}(B_{0})\}$ and by the definition of an inverse, $f^{-1}(B_{0}) = \{x \mid f(x) \in B_{0}\}$.\\
                        For every element $b \in f(f^{-1}(B_{0})), f^{-1}(b) \in f^{-1}(B_{0})$. Therefore (by assigning $x = f^{-1}(b)$), $f(f^{-1}(b)) \in B_{0}$.\\
                        This means $f(f^{-1}(B_{0})) \subset B_{0}$.\\

                        An alternate (and probably better) proof:\\
                        By the definition of an inverse, $f^{-1}(B_{0}) = \{x \mid f(x) \in B_{0}\}$.\\
                        Let $x \in f^{-1}(B_{0})$. By definition, $f(x) \in B_{0}$, so $f^{-1}(B_{0}) \subset B_{0}$.\\

                        Proof that $f(f^{-1}(B_{0})) = B_{0}$ if $f$ is surjective:\\
                        % By the definition of an inverse, $f(f^{-1}(B_{0})) = \{x \mid f^{-1}(x) \in f^{-1}(B_{0})\}$ and $f^{-1}(B_{0}) = \{x \mid f(x) \in B_{0}\}$.\\
                        % % For every $b \in f(f^{-1}(B_{0}))$, there exists some $x \in B_{0}$ such that $f^{-1}(b) = f^{-1}(x)$.\\
                        % $B_{0} \subset B$, so $b \in B_{0} \Rightarrow b \in B$. By the definition of surjectivity, this means that for every $b$ in $B_{0}$, there is some $a \in A_{0}$ such that $b = f(a)$.\\
                        % Every $f(a) \in f(A_{0})$, so every $b \in B_{0}$ is also an element of $f(A_{0})$, so $B_{0} \subset f(A_{0})$.\\
                        % \*Insert a proof that $A_{0} \subset f^{-1}(B_{0})$* IDK, I'm losing braincells.\\\\
                        % Attempt 2:\\
                        Let $b \in B_{0}$. By the definition of surjectivity, there exists an $x$ such that $b = f(x)$. By the definition of an inverse, $f^{-1}(b) = x$. Therefore, $b = f(x) = f(f^{-1}(b))$. Thus, every element $b \in B_{0}$ is equal to an element $f(f^{-1}(b)) \in f(f^{-1}(B_{0}))$, so $B_{0} \subset f(f^{-1}(B_{0}))$. As proven above, $f(f^{-1}(B_{0})) \subset B_{0}$, so $f(f^{-1}(B_{0})) = B_{0}$.

            \end{enumerate}
      \item Let $f:A \to B$ and let $A_{i} \subset A$ and $B_{i} \subset B$ for $i = 0$ and $i = 1$. Show that $f^{-1}$ preserves inclusions, unions, intersections, and differences of sets.
            \begin{enumerate}
                  \item $B_{0} \subset B_{1} \Rightarrow f^{-1}(B_{0}) \subset f^{-1}(B_{1})$\\
                        Let $x \in f^{-1}(B_{0})$. By the definition of an inverse, $f(x) \in B_{0}$. Because $B_{0} \subset B_{1}$, $f(x) \in B_{1}$ as well.\\
                        $f^{-1}(B_{1}) = \{x \mid f(x) \in B_{1}\}$. Thus, every element in $f^{-1}(B_{0})$ is also in $f^{-1}(B_{1})$, so $f^{-1}(B_{0}) \subset f^{-1}(B_{1})$.
                  \item $f^{-1}(B_{0} \cup B_{1}) = f^{-1}(B_{0})\cup f^{-1}(B_{1})$\\
                        Let $x \in f^{-1}(B_{0} \cup B_{1})$. By the definition of an inverse, $f(x) \in B_{0} \cup B_{1}$, so $f(x) \in B_{0} \lor f(x) \in B_{1}$.\\
                        Let $y \in f^{-1}(B_{0})\cup f^{-1}(B_{1})$. By the deinition of inverses, this means $f(y) \in B_{0} \lor f(y) \in B_{1}$.\\
                        Since the definitions for being contained in either set are the same, the two sets are equivalent.
                  \item $f^{-1}(B_{0} \cap B_{1}) = f^{-1}(B_{0})\cap f^{-1}(B_{1})$\\
                        Let $x \in f^{-1}(B_{0} \cap B_{1})$. $f(x) \in B_{0} \cap B_{1}$, so $f(x) \in B_{0} \land f(x) \in B_{1}$. This can be rewritten as $f^{-1}(B_{0})\cap f^{-1}(B_{1})$ and all steps are reversible, showing equality.
                  \item $f^{-1}(B_{0} - B_{1}) = f^{-1}(B_{0}) - f^{-1}(B_{1})$\\
                        Let $x \in f^{-1}(B_{0} - B_{1})$. $f(x) \in B_{0} - B_{1}$, so $x \in B_{0} \land x \notin B_{1}$. This can be rewritten as $f^{-1}(B_{0}) - f^{-1}(B_{1})$ and all steps are reversible, again showing equality.
            \end{enumerate}
            Show that $f$ preserves inclusions and unions only:
            \begin{enumerate}
                  \setcounter{enumii}{4}
                  \item $A_{0} \subset A_{1} \Rightarrow f(A_{0})  \subset f(A_{1})$\\
                        % Let $x \in A_{0}$. $f(x) \in f(A_{0})$. However, because $A_{0} \subset A_{1}$, $x \in A_{1}$, so $f(x) \in f(A_{1})$ as well. Every $f(x) \in f(A_{0})$ corresponds to at least one $x \in A_{0}$, so $f(A_{0})  \subset f(A_{1})$.\\
                        Let $b \in f(A_{0})$. There is some $x \in A_{0}$ such that $f(x) = a$. Because $A_{0} \subset A_{1}$, $x \in A_{1}$, so $f(x) \in f(A_{1})$. Thus, every $a \in f(A_{0})$ is also in $f(A_{1})$, so $f(A_{0}) \subset f(A_{1})$.
                  \item $f(A_{0} \cup A_{1}) = f(A_{0}) \cup f(A_{1})$\\
                        Let $b \in f(A_{0} \cup A_{1})$. There exists some $x \in A_{0} \cup A_{1}$ such that $f(x) = b$. $x \in A_{0} \cup A_{1}$ is logically equivalent to $x \in A_{0} \lor x \in A_{1}$. Thus, $f(x) \in f(A_{0}) \lor f(x) \in A_{1}$, so $f(x) \in f(A_{0}) \cup f(A_{1})$. Substituting $b$ back in for $f(x)$, we get $b \in f(A_{0}) \cup f(A_{1})$. Because every step is reversible, $f(A_{0} \cup A_{1}) = f(A_{0}) \cup f(A_{1})$.
                  \item $f(A_{0} \cap A_{1}) \subset f(A_{0}) \cap f(A_{1})$\\
                        Let $b \in f(A_{0} \cap A_{1})$. There exists some $x \in A_{0} \cap A_{1}$ such that $f(x) = b$. This means that $x \in A_{0} \land x \in A_{1}$, so $f(x) \in f(A_{0}) \land f(x) \in f(A_{1})$, so $f(x) \in f(A_{0}) \cap f(A_{1})$. Thus, every $b \in f(A_{0} \cap A_{1})$ is also in $f(A_{0}) \cap f(A_{1})$, so $f(A_{0} \cap A_{1}) \subset f(A_{0}) \cap f(A_{1})$.\\

                        Show that equality holds if $f$ is injective.\\
                        Let $b \in f(A_{0}) \cap f(A_{1})$. $b \in f(A_{0}) \land b \in f(A_{1})$. There exists an $x_{0} \in A_{0}$ such that $f(x_{0}) = b$ and an $x_{1} \in A_{1}$ such that $f(x_{1}) = b$. Thus, $f(x_{0}) = f(x_{1})$. By the definition of injectivity, $x_{0} = x_{1}$. Thus, $x_{0} \in A_{0} \land x_{0} \in A_{1}$, so $x_{0} \in A_{0} \cap A_{1}$, so $f(x_{0}) \in f(A_{0} \cap A_{1})$. Thus, every $b \in f(A_{0}) \cap f(A_{1})$ is also in $f(A_{0} \cap A_{1})$, so $f(A_{0}) \cap f(A_{1}) \subset f(A_{0} \cap A_{1})$. As shown above, $f(A_{0} \cap A_{1}) \subset f(A_{0}) \cap f(A_{1})$, so $f(A_{0} \cap A_{1}) = f(A_{0}) \cap f(A_{1})$.
                  \item $f(A_{0} - A_{1}) \supset f(A_{0}) - f(A_{1})$\\
                        Let $b \in f(A_{0}) - f(A_{1})$. $b \in f(A_{0}) \land b \notin f(A_{1})$. This means there exists an $x_{0} \in A_{0}$ such that $f(x_{0}) = b$, and there does not exist an $x_{1} \in A_{1}$ such that $f(x_{1}) = b$. Thus, for any $x$ such that $f(x) = b$, $x \in A_{0} \land x \notin A_{1}$, so $x \in A_{0} - A_{1}$. Therefore, every $b \in f(A_{0}) - f(A_{1})$ is also in $f(A_{0} - A_{1})$, so $f(A_{0} - A_{1}) \supset f(A_{0}) - f(A_{1})$.\\

                        Show that equality holds if $f$ is injective.\\
                        Let $b \in f(A_{0} - A_{1})$. There exists an $x \in A_{0} - A_{1}$ such that $f(x) = b$. Thus, $x \in A_{0} \land x \notin A_{1}$. Let there be some $a \in A_{1}$ such that $f(a) = b = f(x)$. However, because $f$ is injective, this implies that $a = x$. $x \notin A_{1}$, so this presents a contradiction. We can then conclude there is no $a \in A_{1}$ such that $f(a) = f(x)$, so $f(x) \notin f(A_{1})$. Thus, $f(x) \in f(A_{0}) \land f(x) \notin f(A_{1})$, so $f(x) \in f(A_{0}) - f(A_{1})$. Therefore, every $b \in f(A_{0} - A_{1})$ is also in $f(A_{0}) - f(A_{1})$, so $f(A_{0}) - f(A_{1}) \supset f(A_{0} - A_{1})$. As shown above, $f(A_{0} - A_{1}) \supset f(A_{0}) - f(A_{1})$, so $f(A_{0} - A_{1}) = f(A_{0}) - f(A_{1})$.
            \end{enumerate}
      \item Show that 2b, 2c, 2f, and 2g of Exercise 2 hold for arbitrary unions and intersections.
            \begin{enumerate}
                  \setcounter{enumii}{1}
                  \item $f^{-1}(\bigcup_{B \in \mathcal{B}} B) = \bigcup_{B \in \mathcal{B}} f^{-1}(B)$\\
                        Let $x \in f^{-1}(\bigcup_{B \in \mathcal{B}} B)$. $f(x) \in \bigcup_{B \in \mathcal{B}} B$, so $f(x) \in B$ for at least one $B \in \mathcal{B}$. Thus, $x \in \bigcup_{B \in \mathcal{B}} f^{-1}(B)$. Every step is reversible, so $f^{-1}(\bigcup_{B \in \mathcal{B}} B) = \bigcup_{B \in \mathcal{B}} f^{-1}(B)$.
                  \item $f^{-1}(\bigcap_{B \in \mathcal{B}} B) = \bigcap_{B \in \mathcal{B}} f^{-1}(B)$\\
                        Let $x \in f^{-1}(\bigcap_{B \in \mathcal{B}} B)$. $f(x) \in \bigcap_{B \in \mathcal{B}} B$, so $f(x) \in B$ for every $B \in \mathcal{B}$. Thus, $x \in \bigcap_{B \in \mathcal{B}} f^{-1}(B)$. Every step is reversible, so $f^{-1}(\bigcap_{B \in \mathcal{B}} B) = \bigcap_{B \in \mathcal{B}} f^{-1}(B)$.
                  \setcounter{enumii}{5}
                  \item $f(\bigcup_{A \in \mathcal{A}} A) = \bigcup_{A \in \mathcal{A}} f(A)$\\
                        Let $b \in f(\bigcup_{A \in \mathcal{A}} A$). There exists some $x \in \bigcup_{A \in \mathcal{A}} A$ such that $f(x) = b$. $x \in A$ for at least one $A \in \mathcal{A}$, so $f(x) = b \in f(A)$ for at least one $A \in \mathcal{A}$. Thus, every $b \in f(\bigcup_{A \in \mathcal{A}} A)$ is also in $\bigcup_{A \in \mathcal{A}} f(A)$, and because every step is reversible, $f(\bigcup_{A \in \mathcal{A}} A) = \bigcup_{A \in \mathcal{A}} f(A)$.
                  \item $f(\bigcap_{A \in \mathcal{A}} A) \subset \bigcap_{A \in \mathcal{A}} f(A)$\\
                        Let $b \in f(\bigcap_{A \in \mathcal{A}} A)$. There exists some $x \in \bigcap_{A \in \mathcal{A}} A$ such that $f(x) = b$. This means $x \in A$ for every $A \in \mathcal{A}$, so  $f(x) = b \in f(A)$ for every $A \in \mathcal{A}$. Thus, every $b \in f(\bigcap_{A \in \mathcal{A}} A)$ is also in  $\bigcap_{A \in \mathcal{A}} f(A)$, so $f(\bigcap_{A \in \mathcal{A}} A) \subset \bigcap_{A \in \mathcal{A}} f(A)$.\\

                        Show that equality holds if $f$ is injective.\\
                        Let $b \in \bigcap_{A \in \mathcal{A}} f(A)$. $b \in f(A)$ for every $A \in \mathcal{A}$. For every $A \in \mathcal{A}$, there exists an $x \in A$ such that $f(x) = b$. Because $f$ is injective, all these $x$'s are equal to each other, so there is a singular $x \in \bigcap_{A \in \mathcal{A}} A$. Thus, $f(x) = b \in f(\bigcap_{A \in \mathcal{A}} A)$. $\bigcap_{A \in \mathcal{A}} f(A) \subset f(\bigcap_{A \in \mathcal{A}} A)$, and because the reverse was proven above, $f(\bigcap_{A \in \mathcal{A}} A) = \bigcap_{A \in \mathcal{A}} f(A)$.
            \end{enumerate}
      \item Let $f: A \to B$ and $g: B \to C$.
            \begin{enumerate}
                  \item If $C_{0} \subset C$, show that $(g \circ f)^{-1}(C_{0}) = f^{-1}(g^{-1}(C_{0}))$.\\
                        $(g \circ f)^{-1}(C_{0}) = \{a \mid (g \circ f)(a) \in C_{0}\} = \{a \mid g(f(a)) \in C_{0}\}$.\\
                        $f^{-1}(g^{-1}(C_{0})) = \{a \mid  f(a) \in g^{-1}(C_{0})\}$ and $g^{-1}(C_{0}) = \{b \mid g(b) \in C_{0}\}$. Therefore, $f^{-1}(g^{-1}(C_{0})) = \{a \mid g(f(a)) \in C_{0}\}$.\\
                        Both $(g \circ f)^{-1}(C_{0})$ and $f^{-1}(g^{-1}(C_{0}))$ share a definition, so they must be equivalent.
                  \item If $f$ and $g$ are injective, show that $g \circ f$ is injective.\\
                        Let $a_{0}$ and $a_{1}$ such that $(g \circ f)(a_{0}) = (g \circ f)(a_{1})$. Thus, $g(f(a_{0})) = g(f(a_{1}))$. Because $g$ is injective, $f(a_{0}) = f(a_{1})$, and because $f$ is injective, $a_{0} = a_{1}$. Thus, $(g \circ f)(a_{0}) = (g \circ f)(a_{1}) \Rightarrow a_{0} = a_{1}$, so $g \circ f$ is injective.
                  \item If $g \circ f$ is injective, what can you say about the injectivity of $f$ and $g$?\\
                        Let $a_{0}$ and $a_{1}$ such that $f(a_{0}) = f(a_{1})$. $g(f(a_{0})) = g(f(a_{1}))$, and because $g \circ f$ is injective, $a_{0} = a_{1}$. Thus, if $g \circ f$ is injective, $f$ is injective as well.\\
                        Nothing can be said about the injectivity of $g$.
                  \item If $f$ and $g$ are surjective, show that $g \circ f$ is surjective.\\
                        $g \circ f : A \to C$. Let $c \in C$. Because $g$ is surjective, there exists a $b \in B$ such that $g(b) = c$. Because $f$ is surjective, there exists an $a \in A$ such that $f(a) = b$. Thus, for every $c \in C$, there exists an $a$ such that $g(f(a)) = c$, so $g \circ f$ is surjective.
                  \item If $g \circ f$ is surjective, what can you say about the surjectivity of $f$ and $g$?\\
                        Let $c \in C$. Because $g \circ f$ is surjective, there is exists $a \in A$ such that $(g \circ f)(a) = g(f(a)) = c$. Let $f(a) = b \in B$; $g(b) = c$. Thus, $c \in C \Rightarrow c = g(b)$ for at least one $b \in B$, so $g$ is surjective.\\
                        Nothing can be said about the surjectivity of $f$.
                  \item $f$ and $g$ are injective $\Rightarrow g \circ f$ is injective.\\
                        $g \circ f$ is injective $\Rightarrow f$ is injective.\\
                        $f$ and $g$ are surjective $\Rightarrow g \circ f$ is surjective.\\
                        $g \circ f$ is surjective $\Rightarrow g$ is surjective.
            \end{enumerate}
      \item Let us denote the \textit{\textbf{identity function}} for a set $C$ by $i_{C}$, i.e. define $i_{C}:C \to C$ to be the function $i_{C}(x) = x$ for all $x \in C$. Given $f: A \to B$, $g:B \to A$ is a \textit{\textbf{left inverse}} of $f$ if $g \circ f = i_{A}$; $h:B \to A$ is a \textit{\textbf{right inverse}} for $f$ if $f \circ h = i_{B}$.
            \begin{enumerate}
                  \item Show that if $f$ has a left inverse, $f$ is injective.\\
                        Let $g$ be the left inverse of $f$. Then, $g(f(x)) = x$. Let $x_{0} \in A$ and $x_{1} \in A$ such that $f(x_{0}) = f(x_{1})$. $g(f(x_{0})) = g(f(x_{1}))$, but also $g(f(x_{0})) = x_{0}$ and $g(f(x_{1})) = x_{1}$. By substituting, we get $x_{1} = x{0}$, so $f(x_{0}) = f(x_{1}) \Rightarrow x_{0} = x_{1}$, so $f$ is injective.\\

                        Show that if $f$ has a right inverse, $f$ is surjective.\\
                        Let $h$ be the right inverse of $f$. Then, $f(h(x)) = x \in B$. Let $a \in A  = h(x)$. Then, $x \in B \Rightarrow x = f(a)$ for at least one $a \in a$, so $f$ is surjective.
                  \item Give an example of a function that has a left inverse but no right inverse.\\
                        The function $f(x) = \sqrt{x}$ has a left inverse $g(x) = x^{2}$, but no right inverse because it is not surjective (there is no value of $x$ such that $f(x) = -1$).
                  \item Give an example of a function that has a right inverse but no left inverse.\\
                        The function $f(x) = x^{2}$ has a right inverse $h(x) = \sqrt{x}$, but no left inverse because it is not injective ($x^{2} = (-x)^{2}$).
                  \item Can a function have more than one left inverse? More than one right inverse?\\
                        A function cannot have more than one left inverse. Let $f:A \to B$ with two left inversesm $g_{1}$ and $g_{2}$. Thus, $g_{1}:B \to A$ and $g_{2}: B \to A$. Let $x \in A$ and let $f(x) = b \in B$. $g_{1}(f(x)) = x = g_{2}(f(x))$, so for every $b \in B$, $g_{1}(b) = g_{2}(b)$. Since $g_{1}$ and $g_{2}$ both have the same domain ($B$), range ($A$), and rule of assignment, $g_{1} = g{2}$. Therefore, no function has 2 or more unique left inverses.

                        A function can have more than one right inverse. $f(x) = x^{2}$ has both $h_{1}(x) = \sqrt{x}$ and $h_{2}(x) = -\sqrt{x}$ as right inverses.
                  \item Show that if $f$ has both a left inverse $g$ and a right inverse $h$, then $f$ is bijective and $g = h = f^{-1}$.\\
                        Because $f$ has a left inverse, $f$ must be injective, and because $f$ has a right inverse $f$ must be surjective (proven in part (a)), meaning $f$ is bijective.\\
                        Because $f$ is bijective, it must have an inverse $f^{-1}:B \to A$ such that $f^{-1}(b) = a$, where $f(a) = b$.\\
                        By substituting, we get that $f^{-1}(f(a)) = a$ for all $a \in A$. This means $f^{-1}$ is a left inverse of $f$, and because a function can have only one left inverse, $f^{-1} = g$.\\
                        We can also substitute to reach $f(f^{-1}(b)) = b$, meaning $f^{-1}$ is a right inverse of $f$. $h:B \to A$ is also a right inverse of $f$, so $f(h(b)) = b$. $f$ is injective, so $h(b) = f^{-1}(b)$ for all $b \in B$, so $h = f^{-1}$.
                        $g = f^{-1}$ and $h = f^{-1}$, so $g = h = f^{-1}$.
            \end{enumerate}
      \item Let $f: \mathbb{R} \to \mathbb{R}$ be the function $f(x) = x^{3} - x$. By restricting the domain and range of $f$ appropriately, obtain from $f$ a bijective function $g$. Draw the graphs of $g$ and $g^{-1}$.\\
            Let $g:[1, \infty) \to [0, \infty)$ be the function $g(x) = x^{3} - x$.\\
            \begin{tabular}{c c}
                  $g(x)$                                & $g^{-1}(x)$                                  \\
                  \includegraphics[scale=0.25]{M1.2.6g} & \includegraphics[scale=0.25]{M1.2.6ginv.png}
            \end{tabular}

\end{enumerate}

\end{document}

### 2023-11-27 Munkres 1.3.tex
Day 34 HW

\documentclass{article}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage{amssymb}
\usepackage{graphicx}
\graphicspath{ {./images/} }

\title{Munkres 1.3}
\date{}
\begin{document}
\maketitle

\begin{enumerate}
    \item Define two points $(x_{0}, y_{0})$ and $(x_{1}, y_{1})$ of the plane to be equivalent if $y_{0} - x_{0}^{2} = y_{1} - x_{1}^{2}$. Check that this is an equivalence relation and describe the equivalence classes.\\
    Reflexivity: Let $(x_{0}, y_{0})$. $y_{0} - x_{0}^{2} = y_{0} - x_{0}^{2}$ so $(x_{0}, y_{0}) \sim (x_{0}, y_{0})$ for every $(x_{0}, y_{0})$.\\
    Symmetry: Let $(x_{0}, y_{0})$ and $(x_{1}, y_{1})$ such that $(x_{0}, y_{0}) \sim (x_{1}, y_{1})$. $y_{0} - x_{0}^{2} = y_{1} - x_{1}^{2}$. Equality is symmetric, so $y_{1} - x_{1}^{2} = y_{0} - x_{0}^{2}$, so $(x_{1}, y_{1}) \sim (x_{0}, y_{0})$. Thus, $(x_{0}, y_{0}) \sim (x_{1}, y_{1}) \Rightarrow (x_{1}, y_{1}) \sim (x_{0}, y_{0})$.\\
    Transitivity: Let $(x_{0}, y_{0})$, $(x_{1}, y_{1})$, and $(x_{2}, y_{2})$ such that $(x_{0}, y_{0}) \sim (x_{1}, y_{1})$ and $(x_{1}, y_{1}) \sim (x_{2}, y_{2})$. $y_{0} - x_{0}^{2} = y_{1} - x_{1}^{2}$ and $y_{1} - x_{1}^{2} = y_{2} - x_{2}^{2}$. By the transitive property of equality, $y_{0} - x_{0}^{2} = y_{2} - x_{2}^{2}$, so $(x_{0}, y_{0}) \sim (x_{2}, y_{2})$. Thus $((x_{0}, y_{0}) \sim (x_{1}, y_{1}) \land (x_{1}, y_{1}) \sim (x_{2}, y_{2})) \Rightarrow (x_{0}, y_{0}) \sim (x_{2}, y_{2})$.
    The relation is satisfies all three properties of an equivalence relation.\\

    An equivalence class of this relation determined by an element $(x_{0}, y_{0})$ is the set of all points $(x, y)$ such that $y-x^{2} = y_{0} - x_{0}^2$. In other words, let $a = y_{0} - x_{0}^2$. For any value of $x$, $(x, x^{2} + a)$ is in the equivalence class. 
    Note that for any point $(x, y)$ in the equivalence class, $(-x, y)$ is also in the equivalence class.
    \item 
\end{enumerate}
\end{document}

### 2023-12-04 Munkres 1.3.tex
Day 35 HW

@@
 By the transitive property of equality, $y_{0} - x_{0}^{2} = y_{2} - x_{2}^{2}$, so $(x_{0}, y_{0}) \sim (x_{2}, y_{2})$.
-Thus $((x_{0}, y_{0}) \sim (x_{1}, y_{1}) \land (x_{1}, y_{1}) \sim (x_{2}, y_{2})) \Rightarrow (x_{0}, y_{0}) \sim (x_{2}, y_{2})$.
+Thus $((x_{0}, y_{0}) \sim (x_{1}, y_{1}) \land (x_{1}, y_{1}) \sim (x_{2}, y_{2})) \Rightarrow (x_{0}, y_{0}) \sim (x_{2}, y_{2})$.\\
             The relation is satisfies all three properties of an equivalence relation.\\
@@
             Note that for any point $(x, y)$ in the equivalence class, $(-x, y)$ is also in the equivalence class.
-    \item 
+      \item Let $C$ be a relation on a set $A$.
+If $A_{0} \subset A$, define the \textit{\textbf{restriction}} of $C$ to $A_{0}$ to be the relation $C \cap (A_{0} \times A_{0})$.
+Show that the restriction of an equivalence relation is an equivalence relation.\\
+            Define $C_{A_{0}}$ to be the restriction of $C$ to $A_{0}$.\\
+            Lemma: $xC_{A_{0}}y \Rightarrow xCy$.\\
+            $xC_{A_{0}}y \iff (x, y) \in C \cap (A_{0} \times A_{0}) \iff (x, y) \in C \land (x, y) \in (A_{0} \times A_{0})$. Thus, by simplifying, $(x, y) \in C \iff xCy$.\\
+
+            Reflexivity: Let $x \in A_{0}$ (and thus also $x \in A$).
+$x \in A$, so $xCx$, so $(x, x) \in C$.
+$x \in A_{0}$, so $(x, x) \in (A_{0} \times A_{0})$.
+$(x, x) \in C \land (x, x) \in (A_{0} \times A_{0})$, so $(x, x) \in C \cap (A_{0} \times A_{0})$, so $xC_{A_{0}}x$ for every $x \in A_{0}$\\
+            Symmetry: Let $x, y \in A_{0}$ such that $xC_{A_{0}}y$.
+$xC_{A_{0}}y \Rightarrow xCy$, so $yCx$ as well.
+Thus $(y, x) \in C$.
+$x \in A_{0} \land y \in A_{0}$, so $(y, x) \in (A_{0} \times A_{0})$ too.
+$(y, x) \in C \cap (A_{0} \times A_{0})$.
+Therefore, $xC_{A_{0}}y \Rightarrow yC_{A_{0}}x$.\\
+            Transitivity: Let $x, y, z \in A_{0}$ such that $xC_{A_{0}}y$ and $yC_{A_{0}}z$.
+$xCy \land yCz$, so $xCz$, so $(x, z) \in C$.
+$x \in A_{0} \land z \in A_{0}$, so $(x, z) \in (A_{0} \times A_{0})$.
+Thus, $(x, z) \in C_{A_{0}}$.
+Therefore, $xC_{A_{0}}y \land yC_{A_{0}}z \Rightarrow xC_{A_{0}}z$.
+      \item Here is a “proof” that every relation $C$ that is both symmetric and transitive is also reflexive: “Since $C$ is symmetric, $aCb$ implies $bCa$.
+Since $C$ is transitive, $aCb$ and $bCa$ together imply $aCa$, as desired.” Find the flaw in this argument.\\
+            The argument simplifies to $aCb \Rightarrow aCa$, meaning there could be a case where $aCb$ and $aCa$ are both false.
+Reflexivity requires $aCa$ to be true in all cases, but this argument depends on $aCb$.
+      \item Let $f:A \to B$ be a surjective function. Let us define a relation on $A$ by setting $a_{0} \sim a_{1}$ if $f(a_{0}) = f(a_{1})$.
+            \begin{enumerate}
+                  \item Show that this is an equivalence relation.\\
+                        Reflexivity: Let $a \in A$. $f(a) = f(a)$ so $a \sim a$ for every $a \in A$.\\
+                        Symmetry: Let $a_{0}, a_{1} \in A$ such that $a_{0} \sim a_{1}$.$f(a_{0}) = f(a_{1})$, so $f(a_{1}) = f(a_{0})$, so $a_{1} \sim a_{0}$.
+Thus, $a_{0} \sim a_{1} \Rightarrow a_{1} \sim a_{0}$.\\
+                        Transitivity: Let $a_{0}, a_{1}, a_{2} \in A$ such that $a_{0} \sim a_{1} \land a_{1} \sim a_{2}$.
+$f(a_{0}) = f(a_{1}) \land f(a_{1}) = f(a_{2})$, so $f(a_{0}) = f(a_{2})$, so $a_{0} \sim a_{2}$.
+Thus, $(a_{0} \sim a_{1} \land a_{1} \sim a_{2}) \Rightarrow a_{0} \sim a_{2}$.
+                  \item Let $A^{*}$ be the set of equivalence classes. Show there is a bijective correspondence of $A^{*}$ with B.\\
+                        Let $x \in A$.
+Let $E = \{y \mid y \sim x\}$ be the equivalence class determined by $x$.
+$E = \{y \mid f(x) = f(y)\}$.
+Let $b = f(x)$; $E = \{y \mid f(y) = b\}$.
+Thus, for every equivalence class in $A^{*}$, there is some $b \in B$ that it corresponds to.
+Because $f$ is surjective, there is some $a \in A$ such that $f(a) = b$ for every $b \in B$, so every $b \in B$ has a corresponding equivalence class (again defined as $E = \{y \mid f(y) = b\}$).
+            \end{enumerate}
 \end{enumerate}

### 2024-01-10 Logic/Munkres 1.3.tex
Move all logic files to subfolder

--- 2023-12-04 Munkres 1.3.tex
@@
 Since $C$ is transitive, $aCb$ and $bCa$ together imply $aCa$, as desired.” Find the flaw in this argument.\\
-            The argument simplifies to $aCb \Rightarrow aCa$, meaning there could be a case where $aCb$ and $aCa$ are both false.
-Reflexivity requires $aCa$ to be true in all cases, but this argument depends on $aCb$.
+            The argument can be rewritten as:\\
+            $aCb \Rightarrow bCa$\\
+            $aCb \land bCa \Rightarrow aCa$\\
+            \rule{5em}{.5pt}\\
+            $aCa$\\
+            Which simplifies to:\\
+            $aCb \Rightarrow aCa$\\
+            \rule{5em}{.5pt}\\
+            $aCa$\\
+            A counterexample can be found in the case where $aCb$ and $aCa$ are both false.\\
+            More specifically: Let $A = \{1, 2\}$.
+Thus, $A \times A = \{(1, 1), (1, 2), (2, 1), (2, 2)\}$.
+Let $C \subset A$ be $\{(1, 1)\}$.
+The premise $2C1 \Rightarrow 2C2$ is true, but $2C2$ is false and $C$ is not reflexive.
+
       \item Let $f:A \to B$ be a surjective function. Let us define a relation on $A$ by setting $a_{0} \sim a_{1}$ if $f(a_{0}) = f(a_{1})$.
@@
                   \item Let $A^{*}$ be the set of equivalence classes. Show there is a bijective correspondence of $A^{*}$ with B.\\
-                        Let $x \in A$.
+                        A conceptual understanding: Let $x \in A$.
 Let $E = \{y \mid y \sim x\}$ be the equivalence class determined by $x$.
@@
 Let $b = f(x)$; $E = \{y \mid f(y) = b\}$.
-Thus, for every equivalence class in $A^{*}$, there is some $b \in B$ that it corresponds to.
-Because $f$ is surjective, there is some $a \in A$ such that $f(a) = b$ for every $b \in B$, so every $b \in B$ has a corresponding equivalence class (again defined as $E = \{y \mid f(y) = b\}$).
+In plain English, every equivalence class is the set of all $a \in A$ that map to a specific $b$ under $f$.\\
+                        Showing injectivity: Let $E_{1}, E_{2} \in A^{*}$ be two equivalence classes that both correspond to $b$.
+This means there is some $a_{1} \in E_{1}$ and some $a_{2} \in E_{2}$ such that $f(a_{1}) = b = f(a_{2})$.
+Thus, $f(a_{1}) = f(a_{2})$, meaning $a_{2} \in E_{1} \land a_{2} \in E_{2}$.
+Distinct equivalence classes are disjoint, but $E_{1}$ and $E_{2}$ overlap, so we can conclude $E_{1} = E_{2}$.
+Thus, $E_{1}$ corresponds to $b$ and $E_{2}$ corresponds to $b$ implies that $E_{1} = E_{2}$, showing that the correspondence from $A^{*}$ to $B$ is injective.\\
+                        Shwoing surjectivity: Let $b \in B$.
+$f$ is surjective, so $f(a) = b$ for at least one $a \in A$.
+Thus, there is a corresponding equivalence class $E = {a \mid f(a) = b} \in A^{*}$ for every $b \in B$.
+The correspondence from $A^{*}$ to $B$ is also surjective.\\
+                        Because the correspondence from $A^{*}$ to $B$ is both injective and surjective, it must be bijective.
             \end{enumerate}

### 2024-01-10 Logic/Test 3 Study Guide.tex
Move all logic files to subfolder

\documentclass{article}
\usepackage[a4paper, margin=.5in]{geometry}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage{graphicx}
\graphicspath{ {./images/} }
\title{Unit 3 Test Cheatsheet}
\date{}
\begin{document}
\maketitle
\textbf{1.1 Set Theory:}\\
Union: $A \cup B = \{x \mid x \in A \lor x \in B\}$; Intersection: $A \cap B = \{x \mid x \in A \land x \in B\}$\\
$A$ and $B$ are disjoint: $A \cap B = \varnothing$\\
Difference: $A - B = \{x \mid x \in A \land x \notin B\}$\\
Distribution: $A \cap (B \cup C) = (A \cap B) \cup (A \cap C)$; $A \cup (B \cap C) = (A \cup B) \cap (A \cup C)$\\
DeMorgan: $A - (B \cup C) = (A - B) \cap (A - C)$; $A - (B \cap C) = (A - B) \cup (A - C)$\\
Power set of A: $\mathcal{P}(A)$ is the set of all subsets of $A$ (including $A$ and $\varnothing$)\\\
Arbitrary union: $\bigcup_{A \in \mathcal{A}} A = \{x \mid x \in A$ for at least one $A \in \mathcal{A}\}$ (where $\mathcal{A}$ is a collection of sets)\\
Arbitrary intersection: $\bigcap_{A \in \mathcal{A}} A = \{x \mid x \in A$ for every $A \in \mathcal{A}\}$ (undefined when $\mathcal{A} = \varnothing$)\\
Cartesian product: $A \times B = \{(a, b) \mid a \in A \land b \in B\}$ (set of all ordered pairs $(a, b)$ where $a \in A$ and $b \in B$)\\

\textbf{1.2 Functions:}\\
Rule of assignment: subset $r$ of the Cartesian product $C \times D$ of two sets, having the property that each element of $C$ appears as the first coordinate of at most one ordered pair in $r$ (i.e. every thing in the domain is mapped to no more than one thing in the range)
\begin{itemize}
    \item $r$ is a rule of assignment if $(c, d) \in r \land (c, d') \in r \Rightarrow d=d'$
    \item Domain: subset of $C$ consisting of all first coordinates of $r$; $\{c \mid$ there exists $d\in D$ such that $(c, d) \in r\}$
    \item Image: subset of $C$ consisting of all second coordinates of $r$ (everything that actually gets mapped to); \\$\{d \mid$ there exists $c\in C$ such that $(c, d) \in r\}$
\end{itemize}
Function $f$: rule of assignment $r$ with a set $B$ that contains the image set of $r$. The domain $A$ of $r$ is also the domain of $f$; image set of $r$ is the image set of $f$; $B$ is the range of $f$.
\begin{itemize}
    \item $f(a)$ is the unique element of $B$ such that $(a, f(a)) \in r$
    \item Let $A_{0} \subset A$. Restriction of $f$ to $A_{0}$: $\{(a, f(a)) \mid a \in A_{0}\}$
    \item Image of $A_{0}$ under $f$: $f(A_{0}) = \{b \mid b = f(a)$ for at least one $a \in A_{0}\}$
    \item Changing the domain or range changes the function
\end{itemize}
Composition of functions: given $f:A \to B$ and $g:B \to C$, $g \circ f:A \to C$ is defined by $g \circ f (a) = g(f(a))$ 
\begin{itemize}
    \item Only defined when the range of $f$ equals the domain of $g$
    \item Alternately: $g \circ f : A \to C = \{(a, c) \mid$ For some $b \in B, f(a) = b \land g(b) = c\}$
\end{itemize}
Injective: no two distinct points in the domain map to the same point in the range; every point in the range is mapped to by at most one point in the domain; $f(a) = f(a') \Rightarrow a = a'$\\
Surjective: every point in the range is mapped to; $b \in B \Rightarrow (b = f(a)$ for at least one $a \in A)$\\
Inverse: exists if $f$ is bijective; $f^{-1}(B_{0}) = \{a \mid f(a) \in B_{0}\}$\\

\textbf{1.3 Relations:}\\
Relation: a relation on $A$ is a subset $C$ of $A \times A$\\
Equivalence relation has 3 properties:
\begin{itemize}
    \item Reflexivity: $xCx$ for every $x \in A$
    \item Symmetry: If $xCy$, then $yCx$
    \item Transitivity: If $xCy$ and $yCz$, then $xCz$
\end{itemize}
Equivalence class determined by x: $\{y \mid y \sim x\}$
\begin{itemize}
    \item Always contains $x$
    \item Two equivalence classes $E$ and $E'$ are either disjoint or equal\\ \includegraphics[scale=0.6]{test3StudyProofEquivalenceClasses.png}
\end{itemize}
Partition of set A: collection of disjoint, nonempty sets of $A$, where the union of these sets is all of $A$; any partition of $A$ comes from exactly one equivalence relation on $A$\\
\includegraphics[scale=0.75]{test3StudyProofPartitions1.png}\\
\includegraphics[scale=0.75]{test3StudyProofPartitions2.png}\\
\end{document}

### 2024-01-17 CPS/LaTeX example.tex
LaTeX Example

\documentclass[10pt]{article}
\usepackage{graphicx}
\usepackage{enumerate} % to get custom labels
\usepackage{hyperref} % for footnotes to link and \hyperref
\usepackage{amsmath} % to get \text in math mode
\usepackage{amssymb} % for things like \mathcal and \mathbb
\usepackage{xcolor} % colored text
\usepackage{fancyhdr}
\usepackage[a4paper, margin=1in]{geometry}

\begin{document}
\begin{center}
    \textbf{\LaTeX{} example}
\end{center}
Hello
\begin{flushright}
    \textit{world}
    \hrule
\end{flushright}
same \hfill line

\begin{enumerate}[(1)] % why is the indentation fucked
    \item words can be:
          \begin{itemize}
              \item \footnotesize{small} (footnote size) \footnote{speaking of footnotes...}
              \item \large{big (Large)}
          \end{itemize}
    \item I can create hyperlinks to websites like \href{https://www.google.com/}{Google} or to other locations in my document (e.g., \hyperlink{a}{“more on that topic later”}).
    \item I use \texttt{array} for stuff like this:
          \begin{center}
            \renewcommand{\arraystretch}{2}
              $\begin{array}{llll} % {llll} says 4 left-aligned columns; why is vertical spacing screwy?
                      \tan{\left( - \dfrac{11 \pi}{6}\right)} & = & \tan{\dfrac{\pi}{6}}                             & \text{since} - \dfrac{11 \pi}{6} \text{ and} \dfrac{\pi}{6} \text{ are coterminal} \\[.8em]
                                                 & = & \dfrac{\sin{\dfrac{\pi}{6}}}{\cos{\dfrac{\pi}{6}}} &                                                                                  \\[.8em]
                                                 & = & \dfrac{\dfrac{1}{2}}{\dfrac{\sqrt{3}}{2}}          &                                                                                  \\[.8em]
                                                 & = & \dfrac{1}{\sqrt{3}}                              & \text{since multiplying by } \dfrac{2}{2} \text{ is legal}
                  \end{array}$
          \end{center}
    \item and \texttt{tabular} for stuff like this:
          \begin{center}
              \begin{tabular}{|l|c|r|}
                  \hline
                  Lorem & ipsum        & dolor \\
                  \S    & $\mathbb{Z}$ & \$    \\
                  \hline
              \end{tabular}
          \end{center}
    \item \textcolor{blue}{second} \quad \textcolor{red}{first}
    \item \includegraphics[scale=.1]{chrispratt.png}
\end{enumerate}
\newpage
\thispagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\hypertarget{a}{Here are those details I promised earlier.}
\end{document}

### 2024-01-22 CPS/ExpandedResume.tex
Expanded Resume

\documentclass{article}
\usepackage[a4paper, margin=0.5in]{geometry}
\usepackage{hyperref}

\begin{document}
\begin{center}
    \LARGE{\textbf{Saffron Liu}} \\
    \normalsize{}
    \href{tel:2106409659}{(210) 640-9659} \textbullet\ \href{mailto:saffron.lu.liu@gmail.com}{saffron.lu.liu@gmail.com}  \textbullet\ \href{https:\\github.com/SapphireHue}{github.com/SapphireHue}
\end{center}
\begin{flushleft}
    \Large{\textbf{Education}} \normalsize
    \rule[10pt]{\textwidth}{0.4pt}\\[-.15in]
    \textbf{Liberal Arts and Science Academy (LASA), Austin, Texas} \hfill \textsl{May 2024}\\
    \textbf{Rank}: 1/293 $\mid$ \textbf{GPA}: 4.7/4.0 $\mid$ \textbf{SAT}: 1570\\
    \textbf{Advanced Coursework}: Linear Algebra, Multivariable Calculus, Logic (Discrete Math), Computational Problem Solving, Advanced Computer Science, Digital Electronics, Latin 5\\
    \textbf{AP Courses (Scores)}: Computer Science A (5), English Language and Composition (4), Latin (5), Physics 1 (5), Physics 2 (5), United States History (5), Calculus BC (5), World History: Modern (5), English Literature and Composition (planned May 2024)

    \vspace{11pt}
    \Large{\textbf{Clubs \& Activities}} \normalsize
    \rule[10pt]{\textwidth}{0.4pt}\\[-.15in]

    \textbf{Intern @ UT Austin Matz Lab} \hfill \textsl{Aug. 2023 - Present}
    \begin{itemize}
        \setlength\itemsep{-0.2em}
        \item Assisting PhD student with projects on population genomics of coral microbes: microbes as indicators of environmental change, microbes for prediction and prevention of coral bleaching
        \item DNA extraction, library preparation (dilution, PCR, gel electrophoresis, barcoding) for next-generation sequencing
        \item Exploratory data analysis using R
    \end{itemize}

    \textbf{CyberPatriot} \hfill \textsl{Aug. 2018 - Present}
    \begin{itemize}
        \setlength\itemsep{-0.2em}
        \item National cybersecurity competition sponsored by the American Airforce Association
        \item Prepare scripts and research to secure Windows and Windows Server machines
        \item Solve forensics challenges
        \item Coordinate team roles and instruct new members annually
    \end{itemize}

    \textbf{Quiz Bowl} \hfill \textsl{Aug. 2021 - Present}
    \begin{itemize}
        \setlength\itemsep{-0.2em}
        \item Hosted intramural tournament (attended by 250+) to recruit new members
        \item Revitalized Science Bowl team
        \item Moderated 2022-2023 and 2023-2024 regional finals for Science Bee, Geography Bee, and History Bee
        \item Facilitated communication between Longhorn and LASA Quiz Bowl to host 2023 Prometheus Science Bowl
        \item 5th place at local tournament, qualifying for the 2024 High School National Championship Tournament
    \end{itemize}

    \textbf{Latin Club}\\
    Praetor (academic manager; elected position) \hfill \textsl{May 2023 - Present}
    \begin{itemize}
        \setlength\itemsep{-0.2em}
        \item Run twice weekly certamen (Latin Jeopardy!) practices and provide players with study resources
        \item Manage club tournament/convention registration
        \item Organized and hosted 2nd annual LASA Certamen tournament (20+ teams from Austin, San Antonio, and Houston)
    \end{itemize}
    Co-President (elected position) \hfill \textsl{Aug. 2022 - May 2023}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Planned and ran weekly club and certamen (Latin Jeopardy) meetings with consistent membership
        \item Managed club tournament/convention registration and convention logistics (travel, lodgings, meals)
        \item Helped host 1st annual LASA Certamen tournament
        \item Promoted the Latin class/club at showcase and outreach events, increasing enrollment and membership (~15/yr)
        \item Brought back annual Broomball fundraiser (\$600) after it stopped due to COVID-19, allowing purchases for the Latin classroom and subsidizing club convention fees
        \item Worked with teacher sponsor to reorganize and more clearly define club roles
    \end{itemize}
    Scribe (elected position) \hfill \textsl{Aug. 2021 - May 2022}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Compiled and distributed officer meeting notes
        \item Helped host 2-day long area convention, overcoming challenges posed by new/unfinished building, new teacher sponsor, and first in-person convention after COVID-19
    \end{itemize}
    Certamen Player / Latin Club Member \hfill \textsl{Jan. 2021 - Present}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Study and mentor younger certamen players in Greek Mythology and Latin Literature year-round
        \item Compete in certamen and academic testing about classical subjects from city to national level
    \end{itemize}

    \newpage
    \textbf{Fundamentals of Computer Science: Teaching Assistant} \hfill \textsl{Aug. 2023 - Present}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Work one-on-one with students to clarify topics and debug code
        \item Answer questions and assist in grading for a class of ~30
    \end{itemize}

    \textbf{LASA Literary Club: Co-founder}\hfill \textsl{Aug. 2020 - May 2022}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Created club to combat social isolation due to COVID
        \item Curated book list and created book groups based on member preferences
        \item Led small discussion groups
    \end{itemize}

    \textbf{Austin Public Library Summer Volunteer} \hfill \textsl{Jun. 2022 - Aug. 2022}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Set up 3 different gaming consoles for the Girl’s Gaming Club
        \item Oversaw group of ~30 people for 4 hours at a time
        \item Commended by program supervisor for independently continuing the program while they were out sick
    \end{itemize}

    \textbf{Piano} \hfill \textsl{2011 - 2020 \& Aug. 2023 - Present}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Enrolled in private lessons from 2011 to 2020; currently enrolled in school independent study piano class
        \item Recent repetoire: \textit{Preludio} and \textit{Melodioso} from \textit{8 Valses Poeticos}, Granados; \textit{Sarabande} and \textit{Gigue} from English Suite No. 2 in A minor, BWV 807, J.S. Bach; \textit{Presto alla tedesca} from Sonata in G, Op. 79, Beethoven; \textit{Nocturne} in G minor, Op. 37 No. 1, Chopin; \textit{Alla turca} from Piano Sonata No. 11, Mozart
    \end{itemize}

    \textbf{Web Development}
    \begin{itemize}
        \setlength\itemsep{-.2em}
        \item Developed \href{https://github.com/ks4729/ks4729.github.io/blob/main/.github/workflows/Spotify_Content.yml}{script} to automatically retrieve user's top songs with Spotify API and update website with Github Actions
        \item Created \href{https://github.com/SapphireHue/Certamen-Buzzer}{website} to parse certamen questions using regular expressions and incrementally display text to users, helping increase access to study materials and combat high barrier to entry
        \item Created \href{https://www.geogebra.org/m/pmvspthz}{GeoGebra file} to create visualization for multivariable calculus class using JavaScript
        \item Projects in progress: database of questions using MongoDB; multiplayer buzzer system using Node.js and Express.js
    \end{itemize}

    \textbf{Other Clubs/Organizations}: Programming in Practice, Women+ in Computer Science, National Honor Society
    \textbf{Other Competitions}: USACO, picoCTF, CyberStart America, Science Bowl, UIL Computer Science, NACLO
    \textbf{Additional Activities}: \href{https://photos.app.goo.gl/tEm2jn1kG2grJVUn6}{crochet}, \href{https://photos.app.goo.gl/ZvdBS4s2AQgb9bm28}{origami}

    \vspace{11pt}
    \Large{\textbf{Education}} \normalsize
    \rule[10pt]{\textwidth}{0.4pt}\\[-.15in]
    \textbf{Academic Awards}\\
    LASA Class of 2024 Valedictorian (2022, 2023)\\
    National Merit Semifinalist (2023)\\
    Rochester Institute of Technology Women in STEM High School Award (2023)\\
    Trustee Award (top 10\% of class) (2023, 2022)\\
    AP Scholar with Distinction (2023)

    \vspace{11pt}
    \textbf{Computer Science}\\
    CyberPatriot: Platinum National Semifinalist (top 7.5\% nationally) (2023)\\
    National Cybersecurity Scholarship Foundation (through CyberStart): Scholar (2023), Finalist (2022)\\
    UIL Computer Science: 2nd place written test and 1st place team at Regionals (2023), 2nd place team at District (2023)\\
    USA Computing Olympiad: Silver Division (2023)\\
    picoCTF: top 10\% (2023)\\

    \vspace{11pt}
    \textbf{Classics}\\
    Nationals 2023: 1st place advanced open certamen, 2nd place Greek mythology, 8th place Latin literature, 6th place academic heptathlon (includes Greco-Roman mythology, Roman history, Greek, and Latin)\\
    State 2023: 1st place advanced competitive certamen, 1st place Greek mythology, 1st place Latin literature\\
    State 2022: 1st place advanced open certamen, 2nd place Greek mythology, 1st place Latin literature\\
    State 2021: 3rd place intermediate competitive certamen, 10th place Greek mythology\\
    Area 2023: 1st place advanced competitive certamen, 1st place Greek mythology, 1st place Latin literature\\
    Area 2022: 2nd place advanced competitive certamen, 1st place Greek mythology, 1st place Latin literature\\
    Area 2021: 1st place intermediate competitive certamen, 2nd place Greek mythology\\

    \vspace{11pt}
    \textbf{Other}\\
    ABRSM Piano Performance Grade 8 (2020)\\
    ABRSM Music Theory Grade 5, passed with merit (2017)\\

\end{flushleft}

\end{document}

### 2024-01-23 CPS/tikz example.tex
Tikz Example - got it to work by reinstalling latex 🥲

\documentclass[11pt]{amsart}
\usepackage[margin=1in]{geometry}            
\usepackage[parfill]{parskip}

% \usepackage{ifplatform, pdftexcmds, catchfile}
% \usepackage{ifplatform, pdftexcmds, catchfile}

\input{Bailey_Barnes_Preamble.sty}

\begin{document}
\begin{center}\textbf{Tikz examples}

\begin{tikzpicture}[scale=1]
    \draw[dashed, gray] (-2,-2)node[below left]{$(-2,-2)$} grid (4,2)node[above right]{$(4,2)$};
    \draw[ultra thick, <->](-1,-1)--(3,1)node[right]{$A$};
    \draw[ultra thick] (-1,1) circle (1);
    \draw[ultra thick, domain=-2:3] plot (\x, {sin(\x r)});
    \draw[ultra thick, domain=0:4, red, <->] plot (\x, {-(pow(\x,2)-4*\x+2)});
\end{tikzpicture}

\begin{tikzpicture}[scale=.5]
    \draw[ultra thick, red, domain=0:1800, samples=500] plot(\x: {\x / 360});
    \draw[ultra thick, blue, domain=0:180,samples=500, fill=blue!15] plot (\x: {6*sin(5*\x)});
    \foreach \x in {1, 2, ..., 6}
        \draw[gray] (0,0) circle (\x);

    \foreach \a in {15, 30, ..., 345}
        \draw[gray] (0:0) -- (\a:6);
    \draw[<->] (-7,0)--(7,0) node[right]{$x$};
    \draw[<->] (0,-7)--(0,7) node[above]{$y$};

\end{tikzpicture}
\end{center}
\end{document}

### 2024-01-24 CPS/Cryptology Packet/GCD.tex
GCD

\documentclass{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{amsmath}
\begin{document}
\begin{flushleft}
      \LARGE\textbf{Greatest Common Divisor}\normalsize\\ \vspace{11pt}

      The \textit{greatest common divisor} (GCD) of two integers, $a$ and $b$, which are not both 0, is the largest integer that divides both $a$ and $b$. We write this gcd($a, b$) and define gcd($0, 0$) to be 0 because otherwise it would be annoying.\\ \vspace{11pt}

      We also have a name for two integers that share no factors. If gcd($a, b$) = 1, we say that $a$ and $b$ are relatively prime.
\end{flushleft}

\begin{enumerate}
      \item Find the greatest common divisor for each of the following pairs of integers.
            \begin{enumerate}
                  \item 15, 35\\
                        5
                  \item 0, 111\\
                        111 (0 is divisible by every integer)
                  \item -12, 18\\
                        6
                  \item 99, 100\\
                        1 (the prime factors of 99 are 3, 3, and 11; the prime factors of 100 are 2, 2, 5, and 5)
                  \item 11, 121\\
                        11
                  \item 100, 102\\
                        2
            \end{enumerate}
      \item Let $a$ be a positive integer. What is gcd($a, 2a$)?\\
            $a$
      \item Let $a$ be a positive integer. What is gcd($a, a^{2}$)?\\
            $a$
      \item Let $a$ be a positive integer. What is gcd($a, a + 1$)?\\
            1. % $\gcd(a, b) \leq \min(a, b)$, so $\gcd(a, a+1) \leq a$. Every positive integer can be divided by 1, so $1 \leq \gcd(a, a+1) \leq a$. Every divisor 
            Every integer can be divided by 1, so $1 \leq \gcd(a, a+1)$. Let $a$ have a divisor $x$ such that $x > 1$. $a\ \text{mod}\ x = 0$, so $(a+1)\ \text{mod}\ x = 1$. Thus, there is no integer greater than 1 that divides both $a$ and $a+1$.
      \item Let $a$ be a positive integer. What is gcd($a, a + 2$)?\\
            If 2 divides $a$, $\gcd(a, a+2) = 2$.\\
            Otherwise, $\gcd(a, a+2) = 1$.
      \item Find the greatest common divisor for each of the following sets of integers.
            \begin{enumerate}
                  \item 8, 10, 12\\
                        2
                  \item 6, 15, 21\\
                        3
                  \item -7, 28, -35\\
                        7
            \end{enumerate}
      \item Find a set of three integers that are mutually relatively prime, but any two of which are not  relatively prime.\\
            6, 15, 10\\
            To achieve this task, I figured I needed 3 numbers, each with 2 distinct prime factors. Each pair of numbers shares a prime factor, so I then reverse-engineered this. The 3 smallest prime numbers are 2, 3, and 5, so I multiplied pairs of these together.
      \item Find four integers that are mutually relatively prime such that any three of these integers are not mutually relatively prime.\\
      105, 70, 42, 30\\ 
      This was found with a similar process. From 4 items, I could construct 4 groups of 3 (each group consists of 1 from the initial group removed), so I wanted 4 unique prime factors; I chose 2, 3, 5, and 7. I constructed 4 groups of 3: \{3, 5, 7\}, \{2, 5, 7\}, \{2, 3, 7\}, and \{2, 3, 5\}. I then took the product of each group.
\end{enumerate}
\end{document}

### 2024-02-05 CPS/Test1Review.tex
Test 1 Review

\documentclass{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{tikz}
\begin{document}
\begin{flushleft}

\textbf{\LaTeX}\\ \vspace{11pt}
\textbf{Functions}\\ \vspace{-11pt}
\begin{verbatim}
To define:
\newcommand{\commandName}[numArgs]{
    function definition here
    use #1 to reference the first argument, #2 to reference the second, etc
}

To use:
\commandName{arg1}{arg2}...
This will output the contents of the function's definition
\end{verbatim}

\textbf{Tikz}\\ \vspace{-11pt}
\begin{tikzpicture}
    \draw[dashed, gray] (-2,-2)node[below left]{$(-2,-2)$} grid (4,2)node[above right]{$(4,2)$};
    \draw[ultra thick, <->](-1,-1)--(3,1)node[right]{$A$};
    \draw[ultra thick] (-1,1) circle (1);
    \draw[ultra thick, domain=-2:3] plot (\x, {sin(\x r)});
    \draw[ultra thick, domain=0:4, red, <->] plot (\x, {-(pow(\x,2)-4*\x+2)});
\end{tikzpicture}

This plot was produced by the following:
\begin{verbatim}
\begin{tikzpicture}
    \draw[dashed, gray] (-2,-2)node[below left]{$(-2,-2)$} grid (4,2)node[above right]{$(4,2)$};
    \draw[ultra thick, <->](-1,-1)--(3,1)node[right]{$A$};
    \draw[ultra thick] (-1,1) circle (1);
    \draw[ultra thick, domain=-2:3] plot (\x, {sin(\x r)});
    \draw[ultra thick, domain=0:4, red, <->] plot (\x, {-(pow(\x,2)-4*\x+2)});
\end{tikzpicture}
\end{verbatim}
\begin{tikzpicture}[scale=.5]
    \draw[ultra thick, red, domain=0:1800, samples=500] plot(\x: {\x / 360});
    \draw[ultra thick, blue, domain=0:180,samples=500, fill=blue!15] plot (\x: {6*sin(5*\x)});
    \foreach \x in {1, 2, ..., 6}
        \draw[gray] (0,0) circle (\x);

    \foreach \a in {15, 30, ..., 345}
        \draw[gray] (0:0) -- (\a:6);
    \draw[<->] (-7,0)--(7,0) node[right]{$x$};
    \draw[<->] (0,-7)--(0,7) node[above]{$y$};

\end{tikzpicture}
\begin{verbatim}
\begin{tikzpicture}[scale=.5]
    \draw[ultra thick, red, domain=0:1800, samples=500] plot(\x: {\x / 360});
    \draw[ultra thick, blue, domain=0:180,samples=500, fill=blue!15] plot (\x: {6*sin(5*\x)});
    \foreach \x in {1, 2, ..., 6}
        \draw[gray] (0,0) circle (\x);

    \foreach \a in {15, 30, ..., 345}
        \draw[gray] (0:0) -- (\a:6);
    \draw[<->] (-7,0)--(7,0) node[right]{$x$};
    \draw[<->] (0,-7)--(0,7) node[above]{$y$};
\end{tikzpicture}
\end{verbatim}

\textbf{Links}\\ \vspace{-11pt}
\begin{verbatim}
\href{target}{text} to do a normal hyperlink
\hyperlink{targetName}{text} to link to other locations in the document; the target name should
    be unique; use it in \hypertarget{targetName}{text}
\footnote{footnote text} to insert a footnote; this will generate the correct number where the
    command is and link it to the text in the footer
\end{verbatim}

\textbf{Miscellaneous}\\ \vspace{-11pt}
\begin{verbatim}
\includegraphics[scale=1]{filename.png} to do an image
\begin{tabular}[p|p|p] where p is either l, r, or c (alignment of the column) 
    and | indicates whether or not there's a vertical line
    - use & to separate columns in a line and \\ to separate lines
    - use \hline to draw a horizontal line
\begin{array}[ppp] works similarly
    - must be in math mode (wrap the array in $)
    - I don't think you can have column separators?
\begin{enumerate} makes a numbered list; \begin{itemize} just gives you bullet points
    - use options with enumerate to get custom numbering (roman numerals, alphabet, etc.) 
        - ex. \begin{enumerate}[(a)] to get (a), (b), (c) and so forth
\textit{} is italics, \textbf{} is bold
\hrule makes a horizontal line across the page
\hfill adds as much horizontal whitespace as needed to make the line's contents take up the
whole page
\frac{top}{bottom} makes a fraction, \dfrac{top}{bottom} gives it more vertical space
\newpage and \pagebreak both create new pages (I Think new page spaces the previous page nicely,
    page break breaks it off as is)
\end{verbatim}
\end{flushleft}

\end{document}

### 2024-02-15 CPS/Cryptology Packet/Character Ciphers.tex
Character Ciphers

\documentclass{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{pythonhighlight}
\usepackage{xcolor}
% \usepackage{listings}
\usepackage{amsmath}
\begin{document}
\begin{flushleft}
    \LARGE\textbf{Character Ciphers}\normalsize\\ \vspace{11pt}
\end{flushleft}
\begin{enumerate}
    \item Using the Caesar Cipher, encrypt the message ATTACK AT DAWN.\\
          I used the following code with $a = 1$ and $b = 3$ to get "DWWDFN DW GDZQ".
          \begin{verbatim}
alphabet = "ABCDEFGHIJKLMNOPQRSTUVWXYZ"
def encrypt(message, a, b):
    out = ""
    for letter in message:
        if letter in alphabet:
            out+=alphabet[(a*(alphabet.find(letter))+b)%26]
        else:
            out+=letter
    return out
          \end{verbatim}
    \item Decrypt the ciphertext message LFDPH LVDZL FRQTX HUHG, which has been encrypted using the Caesar Cipher.\\
          I used the previous \texttt{encrypt()} function with $a = 1$ and $b = -3$ to get "ICAME ISAWI CONQU ERED" (or, "I CAME I SAW I CONQUERED").
    \item Encrypt the message SURRENDER IMMEDIATELY using the affine transformation $C \equiv 11P + 18 \mod{26}$.\\
          \texttt{encrypt("SURRENDER IMMEDIATELY", 11, 18)} produces "IEXXKFZKX CUUKZCSTKJW".
    \item Decrypt the message YLFQX PCRIT, which has been encrypted using the affine transformation $C \equiv 21P + 5 \mod{26}$.
          \begin{verbatim}
alphabet = "ABCDEFGHIJKLMNOPQRSTUVWXYZ"
def multiplicativeInverse(num):
    for i in range(1, 26):
        if (num*i)%26 == 1:
            return i
def decrypt(message, a, b):
    out = ""
    aInverse = multiplicativeInverse(a)
    for letter in message:
        if letter in alphabet:
            out += alphabet[((alphabet.find(letter) - b) * aInverse)%26]
        else:
            out+=letter
    return out
          \end{verbatim}
          \texttt{decrypt("YLFQX PCRIT", 21, 5)} produces "READM YLIPS" ("READ MY LIPS").
    \item If the most common letter in a long ciphertext, encrypted by a shift transformation $C \equiv P + k \mod{26}$ is Q, then what is the most likely value of $k$?\\
          The most common letter in plain English is E, so it's likely that E (4) was encrypted to Q (16). Therefore, $16 \equiv 4 + k \mod{26}$, so algebra tells us $k = 12$.
    \item The message IVQLM IQATQ SMIKP QTLVW VMQAJ MBBMZ BPIVG WCZWE VNZWU KPQVM AMNWZ BCVMK WWSQM was encrypted by a shift transformation $C \equiv P + k \mod{26}$. Use frequencies of letters to determine the value of $k$. What is the plaintext message?\\
          I wrote a script to generate a frequency table:
          \begin{verbatim}
alphabet = "ABCDEFGHIJKLMNOPQRSTUVWXYZ"
frequencies = {}
for letter in alphabet:
    frequencies[letter] = 0

message = input("Message: ")
for letter in message:
    if letter in frequencies:
        frequencies[letter]+=1
print(frequencies)
          \end{verbatim}
          I got the following: \texttt{'A': 3, 'B': 4, 'C': 2, 'D': 0, 'E': 1, 'F': 0, 'G': 1, 'H': 0, 'I': 4, 'J': 1, 'K': 3, 'L': 2, 'M': 9, 'N': 2, 'O': 0, 'P': 3, 'Q': 7, 'R': 0, 'S': 2, 'T': 2, 'U': 1, 'V': 7, 'W': 7, 'X': 0, 'Y': 0, 'Z': 4}. The most common letter is M (9 occurences), followed by Q, V, and W (all 7 occurences).\\
          Thus, I will try to shift E, the most common letter in plaintext, to M, a shift of $k = 8$. This produces "ANIDE AISLI KEACH ILDNO NEISB ETTER THANY OUROW NFROM CHINE SEFOR TUNEC OOKIE" (AN IDEA IS LIKE A CHILD NONE IS BETTER THAN YOUR OWN FROM CHINESE FORTUNE COOKIE).
    \item If the two most common letters in a long ciphertext, encrypted by an affine transformation $C \equiv aP + b \mod{26}$, are W and B, respectively, then what are the most likely values for $a$ and $b$?\\
          The most common letters in English are E and T respectively. Therefore, if we assume E (4) is encrypted as W (22) and T (19) as B (1), we get the following equations: $22 \equiv 4a + b \mod{26}$, $1 \equiv 19a + b \mod{26}$.\\
          We can rearrange the first equation to get $b \equiv 22 - 4a \mod{26}$, then substitute to get $1 \equiv 19a + 22 - 4a \mod{26}$. Further algebra produces $-21 \equiv 15a \mod{26}$, or $5 \equiv 15a \mod{26}$. By testing integers 1 through 26 for $a$, it becomes apparent that $a = 9$.\\
          Substituting that back, we get $b \equiv 22 - 4*9 \equiv 22 - 36 = 12 \mod{26}$. This makes the transformation $C \equiv 9P + 12 \mod{26}$\\
          Now, to verify, because math is difficult. E is 4, $9*4 + 12 = 36 + 12 = 48 \equiv 22 \mod{26}$, so E becomes encrypted as W. T is 19, $9 * 19 + 12 = 183 \equiv 1$, so T becomes encrypted as B. Stuff works out!\\
          This verifies the affine transformation as $C \equiv 9P + 12 \mod{26}$.
    \item The message WEZBF TBBNJ THNBT ADZQE TGTYR BZAJN ANOOZ ATWGN ABOVG FNWZV A was encrypted by an affine transformation $C \equiv aP +b \mod{26}$. The most common letters in the plaintext are A, E, N, and S. What is the plaintext message?\\
          In the ciphertext, the most common letters are A (0), B (1), N (13), and T (19). Therefore, A maps to one of A, B, N, or T, meaning $b$ must be 0, 1, 13, or 19.\\
          This gives us a couple of different cases.\\
          \begin{enumerate}
              \item Case: $b = 0$. This means A maps to A, so E (4) must map to B (1), N (13), or T (19). If E maps to B, then $1 \equiv a*4 \mod 26$. However, this is not possible, because it requires $a*4$ to be odd. The same logic can be used for E mapping to N or T. Therefore, $b$ cannot be $0$.
              \item Case: $b = 1$. This means A maps to B, so E must map to A (0), N (13), or T (19). If E maps to A, then $0 \equiv a * 4 + 1 \mod{26}$. This would again require $a*4$ to be odd, which is not possible. If E maps to N, then $13 \equiv a * 4 + 1 \mod{26}$, so $a = 3$. If E maps to T, then $19 \equiv a*4 + 1$, so $a = 11$. Thus, we have two cases to test: $a = 3, b = 1$ and $a = 11, b = 1$.
                    % a could actually be 3 or 16 (though not actually 16 because of the gcd rule). is there a way to prove its always going to be the smallest one
              \item Case: $b = 13$. This means A maps to N, so E must map to A (0), B (1), or T (19). Again, it's impossible for E to map to A, since that would require $a*4$ to be odd. If E maps to B, $1 \equiv a*4 + 13 \mod{26}$, so $14 \equiv a*4 \mod{26}$, so $a = 14$. If E maps to T, $13 \equiv a*4 + 13 \mod{26}$, so $0 \equiv a*4 \mod{26}$, so $a = 13$. However, because $\gcd(a, 26)$ must be 1, neither of these values of $a$ work, so $b$ cannot be 13.
              \item Case: $b = 19$. This means A maps to T, so E must map to A (0) (impossible because it would require $a*4$ to be odd), B (1), or N (13). If E maps to B, $1 \equiv a*4 +19 \mod{26}$, so $8 \equiv a*4$, so $a = 2 \text{ or } 15$ (by the gcd rule, just 15). If E maps to N, $13 \equiv a*4 + 19 \mod{26}$, so $20 \equiv a*4$, so $a=5$. This gives us two more cases to test: $a = 5, b = 19$ and $a = 15, b = 19$.
          \end{enumerate}
          Now, I take all my cases ($a = 3, b = 1$; $a = 5, b = 19$; $a = 11, b = 1$; $a = 15, b = 19$) and test them. I'm testing the encrypted string, as well as a string of the most common characters in the ciphertext (ABNT), because that should decrypt to some combination of AENS, allowing me to quickly see if the transformation is correct.\\
          None of those worked and I have no idea why $<3$.\\

          I decided to brute force the problem just to see what the correct transformation should be. The most common letters in the ciphertext (A, B, N, and T) each appear 6 times, so I tested every possible value of $a$ (odd numbers 1 through 25, excluding 13) and $b$ (numbers 1 through 25). I filtered the decrypted messages by only outputting messages where A, E, N, and S each appeared exactly 6 times (the most common letters in the plaintext must appear the same number of times as the most common letters in the ciphertext).
          \begin{verbatim}
message = "WEZBF TBBNJ THNBT ADZQE TGTYR BZAJN ANOOZ ATWGN ABOVG FNWZV A"
for a in range(1, 26, 2):
    for b in range(0, 26):
        if(a!=13):
            decrypted = decrypt(message, a, b)
            freqs = freqCount(decrypted)
            if(freqs["A"]==6 and freqs["E"]==6 and freqs["N"]==6 and freqs["S"]==6):
                print(a, b, decrypted)
          \end{verbatim}
          This revealed that the transformation I should have found was $a = 21, b = 13$, and that the plaintext was "THISM ESSAG EWASE NCIPH EREDU SINGA NAFFI NETRA NSFOR MATIO N".\\
          I looked back at my case $b = 13$ to try and figure out where I went wrong. It turns out I did and said a lot of very stupid things. Here is a corrected paragraph, with changes marked:\\
          Case: $b = 13$. This means A maps to N, so E must map to A (0), B (1), or T (19). Again, it's impossible for E to map to A, since that would require $a*4$ to be odd. If E maps to B, $1 \equiv a*4 + 13 \mod{26}$, so $14 \equiv a*4 \mod{26}$, so $a = \colorbox{cyan}{10 \text{ or } 23}$. If E maps to T, $\colorbox{cyan}{19}\equiv a*4 + 13 \mod{26}$, so $6 \equiv a*4 \mod{26}$, so $a = \colorbox{cyan}{8 \text{ or } 21}$. However, because $\gcd(a, 26)$ must be 1, \colorbox{cyan}{$a$ cannot be 10 or 8, so we have two test cases: $a=23, b=13$ and $a = 21, b = 13$.}\\
          The affine transformation used to encode the message was $C \equiv 21P + 13 \mod{26}$ and the message is "THIS MESSAGE WAS ENCIPHERED USING AN AFFINE TRANSFORMATION".

    \item Given two ciphers, plaintext may be encrypted by first using one cipher, and then using another cipher on the result. This procedure produces a \textit{product cipher}. Find the product cipher obtained by using the transformation $C \equiv 5P + 13 \mod{26}$ followed by the transformation $C \equiv 17P + 3 \mod{26}$.\\
          This is essentially a composition of functions. $C \equiv 17(5P + 13) + 3 \mod{26}$, which I can then simplify: $C \equiv 17*5P + 17*13 + 3 \equiv 85P + 221 + 3 \equiv 85P + 224 \mod{26}$. I can then use mod rules to simplify to $C \equiv 7P + 18$.
\end{enumerate}
\end{document}

### 2024-02-15 CPS/Cryptology Packet/Cryptology Scripts/affineShift.py

@@ -1,0 +1,43 @@
+alphabet = "ABCDEFGHIJKLMNOPQRSTUVWXYZ"
+
+def multiplicativeInverse(num):
+    for i in range(1, 26):
+        if (num*i)%26 == 1:
+            return i
+
+def encrypt(message, a, b):
+    out = ""
+    for letter in message:
+        if letter in alphabet:
+            out += alphabet[(a*(alphabet.find(letter))+b)%26]
+        else:
+            out+=letter
+    return out
+
+def decrypt(message, a, b):
+    out = ""
+    aInverse = multiplicativeInverse(a)
+    for letter in message:
+        if letter in alphabet:
+            out += alphabet[((alphabet.find(letter) - b) * aInverse)%26]
+        else:
+            out+=letter
+    return out
+
+def freqCount(message):
+    frequencies = {}
+    for letter in alphabet:
+        frequencies[letter] = 0
+
+    for letter in message:
+        if letter in frequencies:
+            frequencies[letter]+=1
+    return frequencies
+
+message = input("What is your message?")
+a = input("Value of a (leave blank for 1): ")
+a = 1 if len(a)==0 else int(a)
+b = input("Value of b (leave blank for 0): ")
+b = 0 if len(b)==0 else int(b)
+print("Encrypted: " + encrypt(message, a, b))
+print("Decrypted: " + decrypt(message, a, b))

### 2024-02-15 CPS/Cryptology Packet/Cryptology Scripts/charCounter.py

@@ -1,0 +1,10 @@
+alphabet = "ABCDEFGHIJKLMNOPQRSTUVWXYZ"
+frequencies = {}
+for letter in alphabet:
+    frequencies[letter] = 0
+
+message = input("Message: ")
+for letter in message:
+    if letter in frequencies:
+        frequencies[letter]+=1
+print(frequencies)

### 2024-02-28 CPS/Cryptology Packet/Euler's Theorem.tex
Euler's Theorem

\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}

\begin{document}
\begin{flushleft}
    \LARGE\textbf{Euler's Theorem}\normalsize\\ \vspace{11pt}
    First, we’ll define a new function called the Euler phi-function or the Euler totient-function, $\phi(n)$. If n is a positive integer, then $\phi(n)$ represents the number of positive integers not exceeding n that are relatively prime to $n$. You should verify that $\phi(8) = 4$.\\

    \textbf{Theorem 4.1 Euler's Theorem:} If $m$ is a positive integer and $a$ is a positive integer with $\gcd(a, m) = 1$, then $a^{\phi(n)} \equiv 1 \mod{m}$
\end{flushleft}

\begin{enumerate}
    \item Find $\phi(n)$ for the following integers.
        \begin{enumerate}
            \item 7\\
            The following numbers are relatively prime to 7: 1, 2, 3, 4, 5, 6. Therefore, $\phi(7) = 6$.
            \item 10\\
            The following numbers are relatively prime to 10: 1, 3, 7, 9. Therefore, $\phi(10)=4$.
            \item 11\\
            The following numbers are relatively prime to 11: 1, 2, 3, 4, 5, 6, 7, 8, 9, 10. Therefore, $\phi(11) = 10$.
            \item 16\\
            The following numbers are relatively prime to 16: 1, 3, 5, 7, 9, 11, 13, 15. Therefore, $\phi(16) = 8$.
        \end{enumerate}
    \item Find the last digit in the decimal expansion of $3^{1000}$.\\
        We're trying to find $3^{1000} \mod 10$. $\phi(10) = 4$. $3^{1000} = (3^{4})^{250} \equiv 1^{250} = 1 \mod{10}$. The last digit is 1.
    \item Find the last digit in the decimal expansion of $7^{999,999}$.\\
    $7^{999,999} = 7^{1,000,000} \times 7^{-1} = (7^{4})^{250,000} \times 7^{-1}$. We're still in mod 10, so $(7^{4})^{250,000} \times 7^{-1} \equiv 1^{250,000} \times 7^{-1} = 1 \times 7^{-1} = 7^{-1}$. By iterating from numbers 1 to 7, I found that the multiplicative inverse of 7 mod 10 is 3, so the last digit is 3.
    \item Find the number in $\mathbb{Z}_{35}$ congruent to $3^{100,000}$.\\
    The following numbers are relatively prime to 35: 1, 2, 3, 4, 6, 8, 9, 11, 12, 13, 16, 17, 18, 19, 22, 23, 24, 26, 27, 29, 31, 32, 33, 34. Thus, $\phi(35) = 24$. $3^{100,000} = (3^{24})^{4000} \times 3^{4000} \equiv 1^{4000} \times 3^{4000} = 3^{4000} = (3^{24})^{160} \times 3^{160} \equiv 1^{160} \times 3^{160} = (3^{24})^{6} \times 3^{16} \equiv 1^{6} \times 3^{16} = (3^{5})^{3} \times 3 = 243^{3} \times 3 \equiv (-2)^{3} \times 3 = -24 \equiv 11 \mod 35$. Thus, $3^{100,000} \equiv 11 \mod{35}$.
    \item Use Euler’s Theorem to find the multiplicative inverse of 2 modulo 9.\\
    9 is relatively prime to: 1, 2, 4, 5, 7, 8. $\phi(9) = 6$. Therefore, $2^{6} = 2 \times 2^{5} \equiv 1 \mod{9}$. Therefore, the multiplicative inverse of 2 mod 9 is $2^{5}$. $2^{5} = 32 \equiv 5$, so 5 is the multiplicative inverse of 2 modulo 9.
    \item Solve each of the following linear congruences using Euler’s Theorem.
    \begin{enumerate}
        \item $5x \equiv 3 \mod{14}$\\
        $x = 5^{-1} \times 3 \mod{14}$. 14 is relatively prime to: 1, 3, 5, 9, 11, 13. Therefore, $\phi(14) = 6$, so $5^{6} \equiv 1 \mod{14}$, so $5^{-1} \equiv 5^{5} \mod{14}$. $5^{5} = 25 \times 25 \times 5 \equiv -3 \times -3 \times 5 = 45 \equiv 3$. So, $5^{-1} \equiv 3 \mod{14}$, so $x = 3 \times 3 = 9$.
        \item $4x \equiv 7 \mod{15}$\\
        $x = 4^{-1} \times 7 \mod{15}$. 15 is relatively prime to: 1, 2, 4, 7, 8, 11, 13, 14. Therefore, $\phi(15) = 8$, so the multiplicative inverse of 4 modulo 15 is $4^{7}$. $4^{7} = (4^{2})^{3} \times 4 = 16^{3} \times 4 \equiv 1^{3} \times 4 = 4$. Therefore, $x = 4 \times 7 = 28 \equiv 13 \mod{15}$. $x = 13$.
        \item $3x \equiv 5 \mod{16}$\\
        $x = 3^{-1} \times 5 \mod{16}$. 16 is relatively prime to: 1, 3, 5, 7, 9, 11, 13, 15. Therefore, $\phi(16) = 8$, so the multiplicative inverse of 3 modulo 16 is $3^{7}$. $3^{7} = 3^{3} \times 3^{3} \times 3 = 27 \times 27 \times 3 \equiv 11 \times 11 \times 3 = 11 \times 33 \equiv 11 \times 1 = 11$. $x = 11 \times 5 = 55 \equiv 7 \mod{16}$. $x = 7$.
    \end{enumerate}
    \item If $p$ and $q$ are distinct primes, what is $\phi(pq)$? It is safe to assume that $\phi$ is a multiplicative function (i.e., $\phi(pq) = \phi(p) \cdot \phi(q)$) if $p$ and $q$ are distinct primes.\\
    Because $p$ is prime, only 1 and $p$ divide it evenly. Thus, for every number $n$ such that $0 < n < p$, $\gcd(n, p) = 1$. $n$ can be any integer in the range $[1, p-1]$, so there are $p - 1$ numbers less than and relatively prime to $p$, so $\phi(p) = p - 1$. By the same reasoning, $\phi(q) = q - 1$. Since $\phi$ is multiplicative, $\phi(pq) = \phi(p) \cdot \phi(q) = (p - 1)(q - 1)$.
\end{enumerate}
\end{document}

### 2024-03-06 CPS/Projects/Divisors of Zero.tex
Divisors of Zero

\documentclass{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\begin{document}
\begin{flushleft}
    \LARGE\textbf{Divisors of Zero}\normalsize\\ \vspace{11pt}

    Consider the integers modulo $n$. If $n$ is \textbf{not} prime, there is a peculiarity: there exist two non-zero numbers in $\mathbb{Z}_n$ such that their product is 0 mod $n$. For example, $2 \times 3 \equiv 0 \mod{ 6}$. In this example, we would call 2 and 3 \textit{divisors of zero} modulo 6. A divisor of zero, in this context, must be a positive number.\\ \vspace{6pt}

    \textbf{Theorem: } If $p$ is a prime number, then there are no divisors of zero modulo $p$.\\ \vspace{6pt}
    \hrule\vspace{6pt}
    Assume there are two numbers $a$ and $b$ that are divisors of a prime number $p$ modulo $p$. Definitionally, $a, b \in \mathbb{Z}_p$ such that $a \times b \equiv 0 \mod{p}$ and $a \neq 0, b \neq 0$. Using the definition of modulo, we can rewrite this as $k \times p + 0 = a \times b$ where $k \in \mathbb{Z}$. This means that $p$ must evenly divide $ab$. Thus, $p$ must evenly divide either $a$ or $b$.\\
    In the case where $p$ evenly divides $a$, $np = a$ for some positive integer $n$. $a$ must be non-zero, so $n \geq 1$. However, $a$ is must be an integer in $\mathbb{Z}_{p}$, i.e. $\{1, 2, 3, ... p-1 \}$. Therefore there is no $n \geq 1$ where $np = a$, so $p$ cannot evenly divide $a$. The same logic can be used to show that $p$ cannot evenly divide $b$.\\
    We've reached a contradiction: it is not possible for $p$ to evenly divide $a$ or $b$. We can thus conclude the initial assumption of two divisors of zero, $a$ and $b$, modulo $p$, is false. There can be no divisors of zero modulo $p$.
\end{flushleft}
\end{document}

### 2024-03-06 CPS/Cryptology Packet/RSA Cryptosystem.tex
RSA Cryptosystem Part 1

\documentclass{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{graphicx} % Required for inserting images
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\begin{document}
\begin{flushleft}
    \LARGE\textbf{The RSA Cryptosystem}\normalsize\\ \vspace{11pt}
    \textbf{The Theoretical Basis for RSA Encryption}\\
    The RSA algorithm involves 5 numbers: $p, q, E, D,$ and $M$. As a brief introduction, they are:
    \begin{itemize}
        \item two different prime numbers, $p$ and $q$
        \item two numbers in $\mathbb{Z}_{(p-1)(q-1)}$:
              \begin{itemize}
                  \item the "encoding number", $E$, which is relatively prime to $(p-1)(q-1)$
                  \item the "decoding number", $D$, which is the multiplicative inverse of $E$ in $\mathbb{Z}_{(p-1)(q-1)}$
              \end{itemize}
        \item a number in $\mathbb{Z}_pq$ known as the "message number", $M$
    \end{itemize}
\end{flushleft}

\begin{enumerate}
    \item Do the activity \href{https://docs.google.com/spreadsheets/d/1vwZHtPWf8RX4x7-8a8qORGgtC9CZb5UV/edit?usp=sharing&ouid=103144376998261263253&rtpof=true&sd=true}{Modular Inverses}\\
          $E$ must be relatively prime to $(p-1)(q-1)$ in order to have a multiplicative inverse in the system.
    \item Let $p=3$ and $q=5$. Then $(p-1)(q-1) = 8$ and $pq = 15$. Suppose that we choose $E$ to be 3. (We could choose any number that didn’t have a factor of 2 since 2 is the only factor in 8). Find $D$.\\
          $D$ is the multiplicative inverse of $E$ in $\mathbb{Z}_{8}$. Thus, $D = 3$.
    \item Sticking with the same $p, q,$ and $E$ (and therefore the same $D$), complete the table below using the rules of $\mathbb{Z}_{15}$. What do you notice about the entries of the last row?\\
          \begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
              \hline
              $M \mod{pq}$      & 1 & 2 & 3  & 4 & 5 & 6 & 7  & 8 & 9 & 10 & 11 & 12 & 13 & 14 \\
              \hline
              \hline
              $M^{E} \mod{pq}$  & 1 & 8 & 12 & 4 & 5 & 6 & 13 & 2 & 9 & 10 & 11 & 3  & 7  & 14 \\
              \hline
              $M^{ED} \mod{pq}$ & 1 & 2 & 3  & 4 & 5 & 6 & 7  & 8 & 9 & 10 & 11 & 12 & 13 & 14 \\
              \hline
          \end{tabular}\\
          The last row is the same as the first row.
    \item \href{https://docs.google.com/spreadsheets/d/1TyXS6F61i7I05NOypTlqcOyrf-DUD3SJwcPTzKUEWaU/edit?usp=sharing}{Google Sheets RSA Cryptosystem Crux Theorem Activity}\\
          From the activity, it seems that raising a number to a power, then raising that to the multiplicative inverse of the previous power, produces the original number.
\end{enumerate}
\begin{flushleft}
    \textbf{Theorem 5.1 (RSA Cryptosystem Crux):} Suppose that $p$ and $q$ are two distinct prime numbers. Let $E$ be relatively prime to $(p-1)(q-1)$. Then if $D$ is the multiplicative inverse of $E$ in $\mathbb{Z}_{(p-1)(q-1)}$ and if $M$ is any number in $\mathbb{Z}_{pq}$, it will always be that $M^{ED} \equiv M \mod{pq}$.\\

    \textbf{Theorem 5.2:} If $p$ and $q$ are distinct prime numbers and $M$ is a positive integer with $\gcd(M, pq) = 1$, then $M^{(p-1)(q-1)} \equiv 1 \mod{pq}$.\\

    \textbf{Theorem 5.3:} Let $p$ and $q$ be distinct prime numbers, $k$ be a positive integer, and $M$ be a number in $\mathbb{Z}_{pq}$ with $\gcd(M, pq) = 1$. Then $M^{1+k(p-1)(q-1)} \equiv M \mod{pq}$.\\

    \textbf{Theorem 5.4:} Let $p$ and $q$ be distinct primes and $E$ be a number in $\mathbb{Z}_{(p-1)(q-1)}$ such that $E$ is relatively prime to $(p-1)(q-1)$. Then $E$ has a multiplicative inverse in $\mathbb{Z}_{(p-1)(q-1)}$. That is, there exists some $D$ in $\mathbb{Z}_{(p-1)(q-1)}$ such that $D \equiv E^{-1} \mod{(p-1)(q-1)}$.
\end{flushleft}
\begin{enumerate}
    \setcounter{enumi}{4}
    \item Find the primes $p$ and $q$ if $pq = 14,647$ and $\phi(pq) = 14,400$.
          $p$ and $q$ must be distinct, since $pq$ is not a square. Thus, $\phi(pq) = (p-1)(q-1) = pq - p - q + 1 = 14,400$. Thus, we can make a system of equations to solve for $p$ and $q$:\\
          $\left\{
              \begin{array}{lcr}
                  pq             & = & 14,647 \\
                  pq - p - q + 1 & = & 14,400
              \end{array}
              \right.$\\
          Thus, $p + q - 1 = 247$, so $p + q = 248$, so $ p = 248 - q$. We can plug this back into the first equation, so $(248-q)q = 14,647 \to 248q - q^{2} = 14,647 \to 0 = q^{2} - 248q + 14,647$. I put this into the quadratic equation to get $q = 97 \text{ or } 151$. $p$ and $q$ are interchangeable in this equation, so either $p = 151, q = 97$, or $p = 97, q = 151$.
    \item Prove Theorem 5.2 based on what you have already learned (perhaps in a previous section).\\
          Let $p$ and $q$ be distinct prime numbers, and let $M$ be a positive integer with $\gcd(M, pq) = 1$. According to Euler's Theorem, if $m$ is a positive integer and $a$ is a positive integer with $\gcd(a, m) = 1$, then $a^{\phi(m)} \equiv 1 \mod{m}$. By setting $a = M$ and $m = pq$, we get $M^{\phi(pq)} \equiv 1 \mod{pq}$. In the previous section, we also determined that where $p$ and $q$ are distinct primes, $\phi(pq) = (p-1)(q-1)$. Thus, $M^{(p-1)(q-1)} \equiv 1 \mod{pq}$.
    \item Prove Theorem 5.3 based on what you have already learned.\\
          Let $p$ and $q$ be distinct prime numbers, $k$ be a positive integer, and $M$ be a number in $\mathbb{Z}_{pq}$ with $\gcd(M, pq) = 1$. $M^{1+k(p-1)(q-1)} = M \times M^{k(p-1)(q-1)} = M \times (M^{(p-1)(q-1)})^{k} \mod{pq}$. By theorem 5.2, $M^{(p-1)(q-1)} \equiv 1 \mod{pq}$, so $M \times (M^{(p-1)(q-1)})^{k} \equiv M \times 1^{k} \equiv M \mod{pq}$. Thus, $M^{1+k(p-1)(q-1)} \equiv M \mod{pq}$.
    \item Prove Theorem 5.1 (RSA Cryptosystem Crux) based on what you have already learned.\\
          Let $p$ and $q$ be two distinct prime numbers, and let $E$ be relatively prime to $(p-1)(q-1)$. By theorem 5.4, there exists a number $D$ that is the multiplicative inverse of $E$ modulo $(p-1)(q-1)$. This means $ED \equiv 1 \mod{(p-1)(q-1)}$, or $ED = k(p-1)(q-1) + 1$ for some integer $k$. Thus, $M^{ED} = M^{k(p-1)(q-1) + 1} \equiv M \mod{pq}$ by theorem 5.3. Thus, $M^{ED} \equiv M \mod{pq}$.
\end{enumerate}
\end{document}

### 2024-04-16 CPS/Cryptology Packet/RSA Cryptosystem.tex
RSA Cryptosystem Part 2

@@
 Thus, $M^{ED} \equiv M \mod{pq}$.
+\end{enumerate}\hrule \vspace{8pt}   
+\textbf{The RSA Encryption Algorithm}\\ 
+Suppose Alice wants to send Bob a secret message.\\
+Bob picks two prime numbers, $p$ and $q$. Bob also picks an encoding number, $E$, from $Z_{(p-1)(q-1)}$ that is relatively prime to $(p-1)(q-1)$. Bob then makes $pq$ and $E$ public.\\
+Alice converts her number to a message $M$ ($M$ must be less than $pq$, but this can be achieved by breaking the message into segments).
+To encode this message, she calculates $M^{E} \mod{pq}$.
+\vspace{8pt}\\
+For these problems, suppose Bob picks $p = 3, q = 97$. Thus, $pq = 291$ and $(p-1)(q-1) = 192$. He also chooses $E = 5$.
+\begin{enumerate}
+    \item Suppose Alice's message is the number $M = 2$. What number does she send Bob? Describe in your own words how you found this number.\\
+    First, I calculated $M^{E}$. $2^{5} = 32$. $32 \mod{291} = 32$, so she will send Bob $32$.
+    \item Suppose that Alice’s secret message is the number M = 150. What number does she send Bob?\\
+    Again, Alice will send Bob $M^{E} \mod{pq}$, so $150^{5} \mod 291$.
+By modular multiplication rules, $150^{5} \mod{291} = (150^{2})^{2}*150 = 22500^{2} * 150 = (291*77 + 93)^{2} * 150 \equiv 93^2 * 150 = 8649 * 150 \equiv 210 * 150 = 31500 \equiv 72 \mod{291}$.
+Alice sends $72$ to Bob.
+\end{enumerate}
+To decode the message, Bob must find $D$, the multiplicative inverse of $E$ in $\mathbb{Z}_{(p-1)(q-1)}$.
+A computer can easily compute modular inverses with the Euclidean algorithm; in this case $D=77$.
+\begin{enumerate}
+\setcounter{enumi}{2}
+    \item Bob can now decode Alice's message.
+    \begin{enumerate}
+        \item Verify that 77 is indeed the multiplicative inverse of 5 in $\mathbb{Z}_{192}$. Explain in your own words how you know that you are correct.\\
+        To verify, I will find $77 * 5 \mod{192}$. $77*5 = 385 = 192*2 + 1$, so $77 * 5 \equiv 1 \mod{192}$, so $5^{-1} \mod{192}$ is indeed 77.
+        \item Using the RSA Cryptosystem Crux Theorem, explain how Bob can use the number $D$ to decode Alice’s encoded message $M^E$ and recover her original message $M$.\\
+        By the Crux Theorem, $M^{ED} = M \mod{pq}$. The encoded message Bob receives from Alice is $M^{E}$. To find $M$ and decode the message, Bob should calculate $(M^{E})^{D} \mod{pq}$.
+    \end{enumerate}
 \end{enumerate}

### 2024-04-24 CPS/Cryptology Packet/RSA Cryptosystem.tex
RSA Cryptosystme Part 3

@@
     \end{enumerate}
+\end{enumerate} \hrule \vspace{8pt}   
+\textbf{Why RSA Encryption is Useful}
+\begin{enumerate} 
+    \item Suppose that you are listening in on Bob's public transmission and you hear that $pq = 21$ and $E = 5$. Recover $(p - 1)(q - 1)$ and $D$. You have broken Bob's code!\\
+    $pq = 21$, where both $p$ and $q$ are prime.
+Either $p = 1, q = 21$ (which would be stupid), or $p = 3, q = 7$.
+Thus, $(p-1)(q-1) = 2 \times 6 = 12$.
+$D = E^{-1} \mod{12}$, so $D = 5^{-1} \mod{12} = 5$.
+Thus, $(p-1)(q-1) = 12, D = 5$.
+    \item You are a secret agent.
+An evil spy with shallow number theory skills uses the RSA Public Key Coding System in which the public modulus is $n = 1537$, and the encoding exponent is $E = 47$.
+You intercept one of the encoded secret messages being sent to the evil spy, namely the number 570.
+Using your superior number theory skills, decode this message, thereby saving countless people from the fiendish plot of the evil spy.\\
+    By brute-force testing increasing prime numbers, $p = 29, q = 53$.
+Thus, $(p-1)(q-1) = 28 \times 52 = 1456$.
+$D = 31$.
+$M = M^{ED} \mod{pq} = 570^{D} \mod{pq}= 570^{31} \mod{1537} = (570^3)^{10} \times 570 = (120490\cdot1537 - 130)^{10} \times 570 \equiv (-130)^{10} \times 570 = ((-130)^2)^{5} \times 570 \equiv (-7)^{5} \times 570 = -16807 \times 570 \equiv 100 \times 570 = 57000 \equiv 131 \mod{1537}$.
+\end{enumerate}
+These codes are easy to break because the chosen values of $pq$ are relatively small.
+In practice, $pq$s are hundreds of digits long, making them effectively impossible to factor, even for a computer (though choosing numbers too large can make encoding and decoding difficult).\\
+This is a public key encryption algorithm: Bob can make the encoding key, $E$, available to anyone, but only he has the decoding key $D$.
+\begin{enumerate}
+\setcounter{enumi}{2}
+    \item Summarize RSA encryption and the advantages it offers in your own words.\\
+    In RSA encryption, a person publicly releases an encrypting key and a modulus.
+Anyone sending them a message uses those two things to encrypt the message, but only the original person has the decrypting key.
+This allows messages to be encrypted and decrypted without ever needing to transmit a decoding key, eliminating the possibility of interception.
+    \item Can you see any weaknesses in RSA?
+There are a couple and one has to make a few simple modifications to the messages being sent and the algorithm to avoid them.
+For the sake of simplicity, the method discussed here is the simplest version and in practice more sophisticated methods (based on the same idea) are used.\\
+    When implemented correctly, RSA encryption has no weaknesses, barring quantum computers, because being able to efficiently factor large numbers allows for breaking the encryption.\\
+    However, due to a couple of mathematical quirks, the following properties should be ensured:
+    \begin{itemize}
+        \item $p$ and $q$ are both sufficiently large, to ensure a sufficiently large $pq$
+        \item $p-1$ and $q-1$ each have at least one prime factor that is sufficiently large
+        \item $E$ and $D$ are sufficiently large
+    \end{itemize}
+    Other weaknesses of RSA come from its implementation, not the mathematical principles.
 \end{enumerate}

### 2024-05-09 CPS/Cryptology Packet/Extended Euclidean Algorithm.tex
Extended Euclidean Algorithm

\documentclass{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}
\usepackage{minted}
\begin{document}
\begin{flushleft}
    \LARGE\textbf{The Extended Euclidean Algorithm}\normalsize\\ \vspace{11pt} 
    \textbf{Background}\\
    \textbf{Bézout’s Identity}: There exist integers $x, y$ such that $ax + by = \gcd(a, b)$.\\
    Let $a = E$ and $b = (p-1)(q-1)$. In the RSA cryptosystem, $E$ must be relatively prime to $(p-1)(q-1)$. Thus, there exist integers $x, y$ such that $Ex + (p-1)(q-1)y = 1$. In $\mod{(p-1)(q-1)}$: $Ex + (p-1)(q-1)y \equiv 1 \mod{(p-1)(q-1)}$, so $Ex \equiv 1 \mod{(p-1)(q-1)}$, so $x = E^{-1} = D$.\\ \vspace{8pt}

    \textbf{Theorem 6.2}: If $a$ and $b$ are integers such that $a > b > 0$, and $q$ and $r$ are integers such that $a = qb + r$ where $0 \leq r < b$ (Euclidean Division), then $\gcd(a, b) = \gcd(b, r)$. i.e. $\gcd(a, b) = \gcd(a-b, b)$.\\
    The Euclidean Algorithm applies Euclidean Division repeatedly to find the gcd by reducing the problem to smaller pieces.\\\vspace{8pt}

    The Extended Euclidean Algorithm finds the values of $x$ and $y$ specified in Bézout’s Identity by reversing the steps of the Euclidean Algorithm.\\
    Example: Finding $gcd(97, 20)$, and then $x, y$ such that $97x + 20y = gcd(97, 20)$.
    $\begin{array}{|rclclr|rclclcrclr|}
        \hline
         \multicolumn{6}{|c|}{\textbf{The Euclidean Algorithm}}& \multicolumn{10}{|c|}{\textbf{The Extended Euclidean Algorithm}}\\ \hline
          97&=&4\cdot20&+&17&\text{(Step 1)}&1&=&3     &-&1(2)           &=&1(3)  &-&1(2)&\text{(Rearrange step 4)}\\  
          20&=&1\cdot17&+&3 &\text{(Step 2)}& &=&1(3)  &-&1(17-5\cdot3)&=&-1(17)&+&6(3)&\text{(Sub from step 3)}\\
          17&=&5\cdot3 &+&2 &\text{(Step 3)}& &=&-1(17)&+&6(20-1\cdot17)&=&6(20)&-&7(17)&\text{(Sub from step 2)}\\
          3 &=&1\cdot2 &+&1 &\text{(Step 4)}& &=&6(20)&-&7(97-4\cdot20)&=&-7(97)&+&34(20)&\text{(Sub from step 1)}\\
          2 &=&2\cdot1 &+&0 &\text{(Step 5)}&&&&&&&&&&\\
         \hline
    \end{array}$
    Thus, $\gcd(97, 20) = 1$ and $97(-7) + 20(34) = 1$.
\end{flushleft}
\hrule \vspace{8pt}   
\textbf{Problem Set}
    \begin{enumerate}
        \item Use the Euclidean Algorithm to find the greatest common divisor for each of the following pairs of integers without using a computer.
        \begin{enumerate}
            \item 51, 89\\
            $89 = 1\cdot51 + 38 \to 51 = 1\cdot38 + 13 \to 38 = 2\cdot13 + 12 \to 13 = 1\cdot12 + 1 \to 12 = 1\cdot12 + 0$. Thus, $\gcd(51, 89) = \gcd(89, 51) = \gcd(51, 38) = \gcd(38, 13) = \gcd(13, 12) = \gcd(12, 1) = 1$.
            \item 102, 202\\
            $202 = 1\cdot102 + 100 \to 102 = 1\cdot100 + 2 \to 100 = 50\cdot2 + 0$. Thus, $\gcd(102, 202) = 2$.
            \item 666, 1414\\
            $1414 = 2\cdot666 + 82 \to 666 = 8\cdot82 + 10 \to 82 = 8\cdot10 + 2 \to 10 = 5\cdot2 + 0$. Thus, $\gcd(666, 1414) = 2$.
        \end{enumerate}
        \item Use the Extended Euclidean Algorithm to express the greatest common divisor of each of the following pairs of integers as a linear combination of these integers.
        \begin{enumerate}
            \item 51, 89\\
            $1 = 1(13) - 1(12) = 1(13) - 1(38 - 2\cdot13) = -1(38) + 3(13) = -1(38) + 3(51-1\cdot38) = 3(51) - 4(38) = 3(51) - 4(89-1\cdot51) = -4(89) + 7(51)$.
            \item 102, 202\\
            $2 = 1(102) - 1(100) = 1(102) - 1(202 - 1\cdot102) = -1(202) + 2(102)$.
            \item 666, 1414\\
            $2 = 1(82) - 8(10) = 1(82) - 8(666 - 8\cdot82) = -8(666) + 65(82) = -8(666) + 65(1414-2\cdot666) = 65(1414) - 138(666)$.
        \end{enumerate}
        \item Find $50^{-1} \mod{127}$ without using a computer.\\
        By Bézout’s Identity, if $\gcd(50, 127) = 1$, then $50x + 127y = 1$, so $x = 50^{-1} \mod{127}$. \\
        First, we use the Euclidean algorithm to check $gcd(50, 127)$. $127 = 2\cdot50 + 27 \to 50 = 1\cdot27 + 23 \to 27 = 1\cdot23 + 4 \to 23 = 5\cdot4 + 3 \to 4 = 1\cdot3 + 1 \to 3 = 3\cdot1 + 0$. Thus, $gcd(50, 127) = 1$.\\
        Then, we use the Extended Euclidean Algorithm to solve for $x$ and $y$. $1 = 1(4) - 1(3) = 1(4) - 1(23 - 5\cdot4) = -1(23) + 6(4) = -1(23) + 6(27 - 1\cdot23) = 6(27) - 7(23) = 6(27) - 7(50-1\cdot27) = -7(50) + 13(27) = -7(50) + 13(127 - 2\cdot50) = 13(127) - 33(50)$. Thus, $50^{-1}\mod{127} = -33$.
        \item In the RSA cryptosystem, find $D$ given $p = 13$, $q = 17$, and $E = 5$.\\
        $(p-1)(q-1) = 12\cdot16 = 192$. Now, we apply the Extended Euclidean Algorithm on $192, 5$ to find $D = 5{-1} \mod{192}$.\\
        The Euclidean Algorithm: $192 = 38\cdot5 + 2 \to 5 = 2\cdot2 + 1 \to 2 = 2\cdot1 + 0$.\\
        The extension: $1 = 1(5) - 2(2) = 1(5) - 2(192 - 38\cdot5) = -2(192) + 77(5)$. Thus, $D = 77$.
        \item Write a Python function that uses the Euclidean Algorithm to find and return the greatest common divisor of two positive integers. Use your function to find the greatest common divisor for each of the following pairs of integers.
        \begin{minted}{python}
def euclideanGCD(a, b):
    equations = []
    if(a < b):
        a, b = b, a
    r = -1
    while(r!=0):
        q = a // b
        r = a - q * b
        equations.append({'a':a, 'q':q, 'b':b, 'r':r})
        a, b = b, r
    print("gcd(" + str(equations[0]['a']) + ", " + 
        str(equations[0]['b']) + ") = " + str(equations[-1]['b']))
    return equations
        \end{minted}
        \begin{enumerate}
            \item $\gcd(9876543210, 123456789) = 9$
            \item $\gcd(11111111111, 1000000001) = 1$
            \item $\gcd(73433510078091009, 45666020043321) = 3$
        \end{enumerate}
        \item Write a Python function that uses the Extended Euclidean Algorithm to write the greatest common divisor of each of the following pairs of positive integers as a linear combination of those integers.
        \begin{minted}{python}
def extendedAlgorithm(a, b):
    equations = euclideanGCD(a, b)
    g = equations[-1]['b']
    if len(equations) < 2:
        print(str(g) + " = " + str(equations[0]['a']) + "*0 + "
            + str(equations[0]['b']) + "*1") 
        return {'gcd':g, 'a':equations[0]['a'], 'x':0, 'b':equations[0]['b'], 'y':1}
    equations=equations[-2::-1]
    a = 1
    x = equations[0]['a']
    b = -equations[0]['q']
    y = equations[0]['b']
    for i in range(1, len(equations)):
        newA = b
        newX = equations[i]['a']
        newB = a + b * -equations[i]['q']
        newY = x
        a, x, b, y = newA, newX, newB, newY
    print(str(g) + " = " + str(a) + "*" + str(x) + " + " + str(b) + "*" + str(y)) 
    return {'gcd':g, 'a':a, 'x':x, 'b':b, 'y':y}
        \end{minted}
        \begin{enumerate}
            \item $9 = -1371742\cdot9876543210 + 109739361\cdot123456789$
            \item $1 = 99009901\cdot11111111111 + -1100110010\cdot1000000001$
            \item $3 = -6495686882417\cdot73433510078091009 + 10445427205865236\cdot45666020043321$
        \end{enumerate}
    \item In the previous section, we used the Extended Euclidean Algorithm to find that $34 \equiv 20^{-1} \mod{97}$.
    \begin{enumerate}
        \item Find $20^{-1} \mod{97}$ by making clever use of Euler’s Theorem.\\
        By Euler's theorem, $20^{\phi(97)} \equiv 1 \mod{97}$, so $20^{\phi(97)-1} \equiv 20^{-1} \mod{97}$. $97$ is prime, so $\phi(97) = 96$. Thus, $20^{-1} \equiv 20^(96-1) = 20^{95} \mod{97}$. $20^{95} = (20^{2})^{47} \times 20 = 400^{47} \times 20 \equiv 12^{47} \times 20 = (12^{3})^{15} \times 144 \times 20 = 1728^{15} \times 2880 \equiv (-18)^{15} \times -30 = ((-18)^{2})^{7} \times -18 \times -30 = 324^{7} \times 540 \equiv 33^{7} \times 55 = (33^{2})^{3} \times 33 \times 55 = 1089^{3} \times 1815 \equiv 22^{3} \times 69 = 10648 \times 69 \equiv 75 \times 69 = 5175 \equiv 34$.
        \item Explain why using Euler’s Theorem is be a much more computationally difficult strategy than using the Extended Euclidean Algorithm.\\
        Euler's theorem first requires finding the Euler-Totient function of the modulus (let's call it $m$), which in turn requires calculating $m-1$ GCDs. It then involves raising the number you're trying to find the inverse of ($E$) to the power of $m$, which involves either many multiplications and modulo operations, or multiplying very large numbers together.\\
        The Extended Euclidean Algorithm, on the other hand, only needs to calculate a GCD once (and is a calculation that would be performed in calculating $\phi(m)$ for Euler's theorem anyway, since $E < m$). The program uses a little bit more memory to drastically cut down on the number of operations it has to perform.
    \end{enumerate}
    \end{enumerate}
\end{document}

### 2024-05-09 CPS/Cryptology Packet/Cryptology Scripts/extendedEuclideanAlg.py

@@ -1,0 +1,55 @@
+def euclideanGCD(a, b):
+    equations = [] # list of dictionaries of form a = qb + r
+    if(a < b):
+        a, b = b, a
+    r = -1
+    while(r!=0):
+        q = a // b
+        r = a - q * b
+        equations.append({'a':a, 'q':q, 'b':b, 'r':r}) 
+        # print(str(a) + " = " + str(q) + "*" + str(b) + " + " + str(r))
+        a, b = b, r
+    # printDivisionEqs(equations)
+    print("gcd(" + str(equations[0]['a']) + ", " + str(equations[0]['b']) + ") = " + str(equations[-1]['b']))
+    return equations
+
+def printDivisionEqs(equations): 
+    for eq in equations:
+        print(str(eq['a']) + " = " + str(eq['q']) + "*" + str(eq['b']) + " + " + str(eq['r']))
+def printExtendedEq(eq):
+    print(str(eq['gcd']) + " = " + str(eq['a']) + "*" + str(eq['x']) + " + " + str(eq['b']) + "*" + str(eq['y'])) 
+
+def extendedAlgorithm(a, b):
+    equations = euclideanGCD(a, b)
+    g = equations[-1]['b']
+    if len(equations) < 2:
+        printExtendedEq({'gcd':g, 'a':equations[0]['a'], 'x':0, 'b':equations[0]['b'], 'y':1})
+        return {'gcd':g, 'a':equations[0]['a'], 'x':0, 'b':equations[0]['b'], 'y':1}
+    equations=equations[-2::-1]
+    a = 1
+    x = equations[0]['a']
+    b = -equations[0]['q']
+    y = equations[0]['b']
+
+    for i in range(1, len(equations)):
+        # currEquation = {'gcd':g, 'a':a, 'x':x, 'b':b, 'y':y}
+        # printExtendedEq(currEquation)
+
+        newA = b
+        newX = equations[i]['a']
+        newB = a + b * -equations[i]['q']
+        newY = x
+        a, x, b, y = newA, newX, newB, newY
+    currEquation = {'gcd':g, 'a':a, 'x':x, 'b':b, 'y':y}
+    printExtendedEq(currEquation)
+    return currEquation
+
+extendedAlgorithm(2, 10) # testing when one number is the gcd of the other
+print("="*20)
+extendedAlgorithm(20, 97) # testing with example
+print("="*20)
+extendedAlgorithm(9876543210, 123456789) # a
[…]

### 2024-05-10 CPS/Projects/extendedEuclideanAlg.py
Extended Euclidean Algorithm (project edition)

@@ -1,0 +1,40 @@
+def euclideanGCD(a, b):
+    equations = [] # list of dictionaries of form a = qb + r
+    if(a < b):
+        a, b = b, a
+    r = -1
+    while(r!=0):
+        q = a // b
+        r = a - q * b
+        equations.append({'a':a, 'q':q, 'b':b, 'r':r}) 
+        a, b = b, r
+    return equations
+
+def solve(origA, origB, c):
+    equations = euclideanGCD(origA, origB)
+    g = equations[-1]['b']
+    a = x = b = y = 0
+    if c%g != 0:
+        return "no solution"
+    if len(equations) < 2:
+        a = equations[0]['a']
+        x = 0
+        b = equations[0]['b']
+        y = 1
+    else:
+        equations=equations[-2::-1]
+        a = equations[0]['a']
+        x = 1
+        b = equations[0]['b']
+        y = -equations[0]['q']
+
+        for i in range(1, len(equations)):
+            newA = equations[i]['a']
+            newX = y
+            newB = a
+            newY = x + y * -equations[i]['q']
+            a, x, b, y = newA, newX, newB, newY
+    if(a != origA):
+        a, x, b, y = b, y, a, x 
+    factor = c//g
+    return str(c) + " = " + str(a) + "*" + str(x*factor) + " + " + str(b) + "*" + str(y*factor)
